### 2024-03-27 01-Manuscript/Manuscript.qmd
Adding MS file

---
title: "Computational methods in landscape ecology"
bibliography: CLER.bib
---

Landscape ecology is fun [@Turner1989].