### 2024-03-28 01-Manuscript/Manuscript.qmd
Update .gitignore, Project and MS

---
title: "Computational methods in landscape ecology"
author:
  - name: Maximilian H.K. Hesselbarth
    orcid: 0000-0003-1125-9918
    email: mhk.hesselbarth@gmail.com
    affiliations:
      - name: International Institute for Applied Systems Analysis (IIASA) 
        city: Laxenburg
bibliography: CLER.bib
format:
  pdf
---

Landscape ecology is fun [@Turner1989].