### 2024-03-29 04-Various/Outline.qmd
Add outline draft

---
format: html
---

**Journal**: Current Landscape Ecology Reports

**Section**: Methodological Developments in Landscape Ecology

**Article Topic**: Computational methods in landscape ecology

**Editors**: Dr. Dimitri Justeau-Allaire (Section Editor); Dr. Sara Gagné (Editor-in-Chief)

**Due Date**: August 16, 2024

Most recent update: `r format(Sys.time(), "%B %d, %Y // %H:%M:%S")`

# General

1. Please feel free to add topics to the outline. This is just based on some brief brain storming and more a conversation starter
2. Please feel free to add your name to a section as lead. Of course multiple people can work on one section
3. Please feel free to add general comments, remarks, etc.

# Outline

- Introduction to landscape ecology (*Hesselbarth*)
  - Definition of landscape ecology
  - Brief history of landscape ecology
  - Computational aspects of landscape ecology
- Open source vs. proprietary software
- GUI vs. script based software
- Programming language
  - R (*Hesselbarth*, *Nowosad*)
  - Python (*Bosch*)
  - Julia (*Schouten*)
- Classic/traditional tools
- Upcoming/modern tools
  - AI/ML tools
  - Virtual reality
- Spatial planning perspective (*Jung*)
- Spatial patterns (*Nowosad*, *Hesselbarth?*)
- Landscape genetics (*Breitbart*)

# Open Questions

-	Remote sensing?
-	General spatial data developments?

# Literature to look at
[10.1007/s10980-021-01366-9](www.doi.org/10.1007/s10980-021-01366-9)