### 2024-03-29 04-Various/Authors.qmd
Update authors

---
title: Authors
format: 
  html:
    page-layout: full
---

Table with authors information (in random order)

| Name | Affiliation | E-Mail | ORCID |
|:------|-------------|--------|-------:|
| Jakub Nowosad | Adam Mickiewicz Uni | nowosad.jakub@gmail.com | 0000-0002-1057-3721 |
| Martin Jung | IIASA | jung@iiasa.ac.at | 0000-0002-7569-1390 |
| Marti Bosch | EPFL | marti.bosch@epfl.ch | 0000-0001-8735-9144 |
| Rafael Schouten | University of Copenhagen | rafael.schouten@sund.ku.dk | 0000-0002-8380-0884 |
| Sophie Breitbart | University of Toronto | sophie.breitbart@mail.utoronto.ca | 0000-0001-9641-9786 |
| Maximilian Hesselbarth | IIASA | mhk.hesselbarth@gmail.com | 0000-0003-1125-9918 |