### 2024-03-29 01-Manuscript/Manuscript.qmd
Adding MS

---
title: "Computational methods in landscape ecology"
author:
  - name: Maximilian H.K. Hesselbarth
    orcid: 0000-0003-1125-9918
    email: mhk.hesselbarth@gmail.com
    affiliations:
      - name: International Institute for Applied Systems Analysis (IIASA) 
        city: Laxenburg
bibliography: CLER.bib
format:
  pdf:
    documentclass: scrartcl
    number-sections: true
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
---

# Introduction

Landscape ecology is fun [@Turner1989].
There are many things to explore [@With2019].

\printbibliography