### 2024-04-10 04-Various/Outline.qmd
Update outline

---
title: Outline
format: html
---

**Journal**: Current Landscape Ecology Reports

**Section**: Methodological Developments in Landscape Ecology

**Article Topic**: Computational methods in landscape ecology

**Editors**: Dr. Dimitri Justeau-Allaire (Section Editor); Dr. Sara Gagné (Editor-in-Chief)

**Due Date**: August 16, 2024

Most recent update: `r format(Sys.time(), "%B %d, %Y // %H:%M:%S")`

# General

1. Please feel free to add topics to the outline. This is just based on some brief brain storming and more a conversation starter
2. Please feel free to add your name to a section as lead. Of course multiple people can work on one section
3. Please feel free to add general comments, remarks, etc. (using your initals)
4. Knit and push changes

# Outline

- Introduction to landscape ecology (*Hesselbarth*)
  - Brief definition of landscape ecology
  - Computational aspects of landscape ecology
- Software and programming language for landscape ecology
  - Open source vs. proprietary software
  - GUI vs. script based software
  - Programming language
    - R (*Hesselbarth*, *Nowosad*)
    - Python (*Bosch*)
    - Julia (*Schouten*)
- Computational methods
  - Visualization (i.e., maps)
  - Spatial patterns (*Nowosad*, *Hesselbarth?*)
  - Connectivity
  - Simulations (e.g., neutral landscapes) (*Hesselbarth*)
  - Optimization
  - Spatial planning perspective (*Jung*)
  - Landscape genetics (*Breitbart?*)
- Discussion
  - General spatial data developments
  - AI impacts on landscape ecology

# Open Questions

-	Remote sensing developments?