### 2024-04-10 04-Various/Literature.qmd
Update literature

---
title: Interesting literature
format: html
---

Most recent update: `r format(Sys.time(), "%B %d, %Y // %H:%M:%S")`

# Literature to look at
[10.1007/s10980-021-01366-9](www.doi.org/10.1007/s10980-021-01366-9)
[https://doi.org/10.1111/2041-210X.14061](https://doi.org/10.1111/2041-210X.14061)