### 2024-05-15 Sections/introduction.qmd
Starting to move things around

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth -->

# Introduction

Landscape ecology is fun [@Turner1989].
There are many things to explore [@With2019].