### 2024-05-17 Sections/simulations.qmd
Add new MS and Sections structure

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth, Schouten -->

# Simulation models

## Neutral landscapes
<!-- Hesselbarth -->

## Process-based simulation
<!-- Schouten, Hesselbarth -->