### 2024-05-17 Sections/software.qmd

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth, Nowosad, Bosch, Jung, Schouten  -->

# Software and programming languages

## Open source vs. proprietary software

## GUI vs. script-based software

## Programing languages

### R
<!-- Hesselbarth -->

### Python
<!-- Bosch -->

### Julia
<!-- Schouten -->