### 2024-05-17 Sections/introduction.qmd
Start to work on intro

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth -->

# Introduction

Landscapes are typically defined as mosaics of different land covers, habitats, ecosystems, or land use systems [@Forman1986,@Forman1995] - with empathize on existing heterogeneity in at least one factor of interest [@Turner2015].
The link of spatial heterogeneity and abiotic as well as biotic processes is the fundamental concept of landscape ecology.
This includes potential feedbacks between processes and heterogeneity, spatial and temporal flows across heterogenous landscapes, or the management of heterogeneity in landscapes [@Risser1984,@Turner1989].
However, due to the spatial context of heterogeneity, the pattern-process link is scale dependent [@Levin1992,@Wiens1989,@Wu2013].
<!-- MH: Add remote sensing data paper -->
Because landscape ecology is a cross-disciplinary field including, e.g., social and political sciences, geography, or ecology and evolution [@Wiens1997,With2019] and the general increase of available data [@Jaric2020,@Nathan2022,@Wuest2020], there is the need for sophisticated computational methods.

<!-- Computational methods refer to techniques and procedures used to solve problems and perform calculations using computers -->
<!-- Computational methods in landscape ecology refer to the use of computer-based algorithms and models to analyze and understand the patterns and processes of landscapes. -->