### 2024-05-22 Manuscript.tex
Minor changes intro

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  a4paperpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[margin=25mm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{lineno}\linenumbers
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Computational methods in landscape ecology},
  pdfauthor={Maximilian H.K. Hesselbarth},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Computational methods in landscape ecology}

  \author{Maximilian H.K. Hesselbarth}
            \affil{%
                  International Institute for Applied Systems Analysis,
                  Biodiversity, Ecology, and Conservation Group,
                  Laxenburg (Austria)
              }

\date{}
\begin{document}
\maketitle

\setstretch{1.5}
\section{Introduction}\label{introduction}

Landscapes are typically defined as mosaics of different land covers,
habitats, ecosystems, or land use systems (Forman and Godron 1986;
Forman 1995) - with empathize on heterogeneity of at least one factor of
interest (Turner and Gardner 2015). Linking the spatial heterogeneity
and ecological processes, including potential feedbacks, is the
fundamental concept of landscape ecology (Risser et al. 1984; Turner
1989, 2005). Typical research topics include ecological flows in
landscape mosaics, drivers and consequences of land use and land cover
change, nonlinear dynamics and landscape complexity, scale and scaling,
spatial analysis and landscape modeling, linking landscape metrics to
ecological processes, integrating humans and their activities into
landscape ecology, landscape conservation and sustainability, or
accuracy assessment and uncertainty analysis (Wu and Hobbs 2002; Wu
2013).

Computational science involves analyzing abstracted core mechanisms of
research questions using data and algorithms. Thus, computational
ecology can be defined as computational science that is used to address
ecological research questions with focus on complex, adaptive systems
and data-driven approaches (Poisot et al. 2019). Computational ecology
in general and simulations specifically are of importance for ecology
because ecological data is often context- and scale-dependent and
unfeasible to study in reproducible, replicable, and controllable
experimental settings (Petrovskii and Petrovskaya 2012; but see Wiersma
2022). Because landscape ecology is a cross-disciplinary field
including, e.g., social sciences, geography, or ecology and evolution
(Wiens 1997; With 2019), the general increase of available data (Chi et
al. 2016; Jarić et al. 2020; Wüest et al. 2020; Nathan et al. 2022), or
the complex landscape systems (Newman et al. 2019), there is the need
for sophisticated computational methods in order to link patterns and
processes. Thus, computational methods are one of the most important
tools of modern scientific research (Prlić and Procter 2012; Wilson et
al. 2014).

This paper aims to introduce the latest developments of computation
methods in landscape ecology. However, it is neither a systematic review
of the literature nor general introduction to (computational) landscape
ecology. For a more comprehensive introductions to landscape ecology in
general, please see Turner and Gardner (2015) or Gergel and Turner
(2017).

\section{Spatial patterns}\label{spatial-patterns}

Landscape ecology is fun (Turner 1989). There are many things to explore
(With 2019).

\subsection{Landscape metrics}\label{landscape-metrics}

\subsection{Vector metrics}\label{vector-metrics}

\subsection{Surface metrics}\label{surface-metrics}

\subsection{Entropy}\label{entropy}

\subsection{Landscape Mosaic}\label{landscape-mosaic}

\section{Spatial planning}\label{spatial-planning}

Landscape ecology is fun (Turner 1989). There are many things to explore
(With 2019).

\newpage{}

\subsubsection{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-Chi2016}
Chi M, Plaza A, Benediktsson JA, et al (2016) Big data for remote
sensing: Challenges and opportunities. Proceedings of the IEEE
104:2207--2219. \url{https://doi.org/10.1109/JPROC.2016.2598228}

\bibitem[\citeproctext]{ref-Forman1995}
Forman RTT (1995) Land mosaics: The ecology of landscapes and regions.
Cambridge University Press, Cambridge, UK

\bibitem[\citeproctext]{ref-Forman1986}
Forman RTT, Godron M (1986) Landscape ecology. Wiley \& Sons,
Chichester, UK

\bibitem[\citeproctext]{ref-Gergel2017}
Gergel SE, Turner MG (eds) (2017)
\href{https://doi.org/10.1007/978-1-4939-6374-4}{Learning landscape
ecology}. Springer New York, New York, NY

\bibitem[\citeproctext]{ref-Jaric2020}
Jarić I, Correia RA, Brook BW, et al (2020) iEcology: Harnessing large
online resources to generate ecological insights. Trends in Ecology \&
Evolution 35:630--639. \url{https://doi.org/10.1016/j.tree.2020.03.003}

\bibitem[\citeproctext]{ref-Nathan2022}
Nathan R, Monk CT, Arlinghaus R, et al (2022) Big-data approaches lead
to an increased understanding of the ecology of animal movement. Science
375:eabg1780. \url{https://doi.org/10.1126/science.abg1780}

\bibitem[\citeproctext]{ref-Newman2019}
Newman EA, Kennedy MC, Falk DA, McKenzie D (2019) Scaling and complexity
in landscape ecology. Frontiers in Ecology and Evolution 7:293.
\url{https://doi.org/10.3389/fevo.2019.00293}

\bibitem[\citeproctext]{ref-Petrovskii2012}
Petrovskii S, Petrovskaya N (2012) Computational ecology as an emerging
science. Interface Focus 2:241--254.
\url{https://doi.org/10.1098/rsfs.2011.0083}

\bibitem[\citeproctext]{ref-Poisot2019}
Poisot T, LaBrie R, Larson E, et al (2019) Data-based, synthesis-driven:
Setting the agenda for computational ecology. Ideas in Ecology and
Evolution 12: \url{https://doi.org/10.24908/iee.2019.12.2.e}

\bibitem[\citeproctext]{ref-Prlic2012}
Prlić A, Procter JB (2012) Ten simple rules for the open development of
scientific software. PLoS Computational Biology 8:e1002802.
\url{https://doi.org/10.1371/journal.pcbi.1002802}

\bibitem[\citeproctext]{ref-Risser1984}
Risser PG, Karr JR, Forman RTT (1984) Landscape ecology: Directions and
approaches. Illinois Natural History Survery Special Publication 2:7--14

\bibitem[\citeproctext]{ref-Turner1989}
Turner MG (1989) Landscape ecology: The effect of pattern on process.
Annual Review of Ecology and Systematics 20:171--197.
\url{https://doi.org/10.1146/annurev.es.20.110189.001131}

\bibitem[\citeproctext]{ref-Turner2005}
Turner MG (2005) Landscape ecology: What is the state of the science?
Annual Review of Ecology, Evolution, and Systematics 36:319--344.
\url{https://doi.org/10.1146/annurev.ecolsys.36.102003.152614}

\bibitem[\citeproctext]{ref-Turner2015}
Turner MG, Gardner RH (2015) Landscape ecology in theory and practice:
Pattern and process, 2nd edition. Springer, New York

\bibitem[\citeproctext]{ref-Wiens1997}
Wiens JA (1997)
\href{https://doi.org/10.1016/B978-012323445-2/50005-5}{Metapopulation
dynamics and landscape ecology}. In: Metapopulation biology. Elsevier,
pp 43--62

\bibitem[\citeproctext]{ref-Wiersma2022}
Wiersma YF (2022) A review of landscape ecology experiments to
understand ecological processes. Ecological Processes 11:57.
\url{https://doi.org/10.1186/s13717-022-00401-0}

\bibitem[\citeproctext]{ref-Wilson2014}
Wilson G, Aruliah DA, Brown CT, et al (2014) Best practices for
scientific computing. PLoS Biology 12:e1001745.
\url{https://doi.org/10.1371/journal.pbio.1001745}

\bibitem[\citeproctext]{ref-With2019}
With KA (2019) Essentials of landscape ecology, 1st edn. Oxford
University Press, Oxford, UK

\bibitem[\citeproctext]{ref-Wu2013}
Wu J (2013) Key concepts and research topics in landscape ecology
revisited: 30 years after the allerton park workshop. Landscape Ecology
28:1--11. \url{https://doi.org/10.1007/s10980-012-9836-y}

\bibitem[\citeproctext]{ref-Wu2002}
Wu J, Hobbs R (2002) Key issues and research priorities in landscape
ecology: An idiosyncratic synthesis. Landscape Ecology 17:355--365.
\url{https://doi.org/10.1023/A:1020561630963}

\bibitem[\citeproctext]{ref-Wuest2020}
Wüest RO, Zimmermann NE, Zurell D, et al (2020) Macroecology in the age
of big data -- where to go from here? Journal of Biogeography 47:1--12.
\url{https://doi.org/10.1111/jbi.13633}

\end{CSLReferences}

\end{document}