### 2024-06-01 Sections/software.qmd
adds new ideas

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth, Nowosad, Bosch, Jung, Schouten  -->

<!-- "The future applications of landscape pattern metrics will almost certainly depend on the software and analysis tools available" -- @costanzaDescribingAnalyzingLandscape2019 -->

# Software and programming languages

## Open source vs. proprietary software

## GUI vs. script-based software

## Programing languages

### R
<!-- Hesselbarth -->
landscapemetrics
geodiv
vectormetrics
bespatial
LandComp
motif
?GLCMTextures
?rasterdiv
?terra

### Python
<!-- Bosch -->

### Julia
<!-- Schouten -->

## Other open source software

GuidosToolbox
?GRASS GIS
?QGIS: LecoS plugin