### 2024-06-19 Manuscript.tex
Some updates on intro and spatial-patterns

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  a4paperpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[margin=25mm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{lineno}\linenumbers
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Computational methods in landscape ecology},
  pdfauthor={Maximilian H.K. Hesselbarth; Jakub Nowosad; Marti Bosch; Martin Jung; Rafael Schouten},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Computational methods in landscape ecology}

  \author{Maximilian H.K. Hesselbarth}
            \affil{%
                  International Institute for Applied Systems Analysis,
                  Biodiversity, Ecology, and Conservation Group,
                  Laxenburg, Austria
              }
        \author{Jakub Nowosad}
            \affil{%
                  Adam Mickiewicz University, Poznan, Poland
              }
        \author{Marti Bosch}
            \affil{%
                  École polytechnique fédérale de Lausanne, Lausanne,
                  Switzerland
              }
        \author{Martin Jung}
            \affil{%
                  International Institute for Applied Systems Analysis,
                  Biodiversity, Ecology, and Conservation Group,
                  Laxenburg, Austria
              }
        \author{Rafael Schouten}
            \affil{%
                  Globe Institute, University of Copenhagen, Copenhagen,
                  Denmark
              }

\date{}
\begin{document}
\maketitle

\setstretch{1.5}
\section{Introduction}\label{introduction}

Landscapes are typically defined as mosaics of different land covers,
habitats, ecosystems, or land use systems (Forman and Godron 1986;
Forman 1995) - with empathize on heterogeneity of at least one factor of
interest (Turner and Gardner 2015). Linking spatial heterogeneity and
ecological processes, including potential feedbacks, is the fundamental
concept of landscape ecology (Risser et al. 1984; Turner 1989, 2005).
Typical research topics include ecological flows in landscape mosaics,
drivers and consequences of land use and land cover change, nonlinear
dynamics and landscape complexity, scale and scaling, spatial analysis
and landscape modeling, linking landscape metrics to ecological
processes, integrating humans and their activities into landscape
ecology, landscape conservation and sustainability, or accuracy
assessment and uncertainty analysis (Wu and Hobbs 2002; Wu 2013).

Computational science involves analyzing abstracted core mechanisms of
research questions using data and algorithms. Thus, computational
ecology can be defined as computational science that is used to address
ecological research questions with focus on complex, adaptive systems
and data-driven approaches (Poisot et al. 2019). Computational ecology
in general and simulations specifically are of importance for ecology
because ecological data is often context- and scale-dependent and
unfeasible to study in controllable, reproducible, and replicable
experimental settings (Petrovskii and Petrovskaya 2012; but see Wiersma
2022). Because landscape ecology is a cross-disciplinary field
including, e.g., social sciences, geography, or ecology and evolution
(Wiens 1997; With 2019), the general increase of available data (Chi et
al. 2016; Jarić et al. 2020; Wüest et al. 2020; Nathan et al. 2022), or
the complex landscape systems (Newman et al. 2019), there is the need
for sophisticated computational methods in order to link heterogeneity
patterns and ecological processes. Thus, computational methods are one
of the most important tools of modern scientific research (Prlić and
Procter 2012; Wilson et al. 2014).

Here, we aim to introduce the latest developments of computational
methods in landscape ecology. However, we do not provide a systematic
literature review, or a general introduction to (computational)
landscape ecology. For a more comprehensive, general introductions to
landscape ecology, please see Turner and Gardner (2015), or Gergel and
Turner (2017). First, we will introduce recent computational
developments of the most important topics of landscape ecology. This
includes, for example, spatial patterns quantification, connectivity and
movement, or model simulations. Second, we will introduce common
software tools that implement or are potentially capable of implementing
these developments. However, due to its many advantages, we will focus
on open-source software (von Krogh and von Hippel 2006), and
specifically scripting languages, i.e., R, Python, and Julia. Last, we
will discuss potential future needs developments of computational
methods in landscape ecology.

\section{Spatial patterns}\label{spatial-patterns}

Spatial patterns can be defined as the scale-dependent predictability of
the physical arrangement of observations (Dale 1999), or as clearly
identifiable structures in nature itself or data extracted from nature
(Grimm et al. 1996). Importantly, patterns observed in nature contain
information about the history of the system, such as demographic
processes, dispersal characteristics, or climatic patterns (Wiegand et
al. 2003). However, spatial patterns are not only a result of processes,
but could also be a driver of them. Thus, untangling the history of a
landscape from its spatial patterns is one of the core concepts of
landscape ecology.

In order to analyze spatial patterns they must be translated into
simplified and quantifiable representations of the natural world. Thus,
the choice of the data types, spatial and temporal scales, or the
thematic resolution are crucial for the analysis. Potential data types
include raster or vector data to represent landscape characteristics
using grid cells or points, lines and polygons, respectively. For the
spatial scale the most important features are the resolution and extent.
The increasing availability of temporal time series data allows analysis
of spatial patterns over time, introducing an additional temporal scale.
For the thematic resolution, which often refers to the number or
definitions of the categories (e.g, LULC classes), many classification
systems exists, such as the NLCD, the CORINE, or the FAO systems (Nedd
et al. 2021).

\subsection{Landscape metrics}\label{landscape-metrics}

Traditionally and still the most prominent, computational approaches to
spatial patterns in landscape ecology revolve around the patch mosaic
model, where the landscape is represented as raster data using
categorical raster data (Costanza et al. 2019).

\begin{itemize}
\tightlist
\item
  Quantify composition and configuration
\item
  Widely applied since FRAGSTATS release
\item
  Several alternative tools available and still widely used
\end{itemize}

\subsection{Surface metrics}\label{surface-metrics}

\href{https://doi.org/10.1007/s10980-018-0685-1}{Surface metrics}

\begin{itemize}
\tightlist
\item
  More recent development
\item
  Quantify continuous raster data (gradient surface models)
\item
  More realistic representation of landscapes (?)
\item
  Based on molecular physics/medicine
\item
  Not widely adopted yet (due to missing tools?)
\end{itemize}

\subsection{Entropy}\label{entropy}

Entropy in landscape ecology is mainly used to quantify the complexity
of the landscape (spatial heterogeneity), and less often
unpredictability (temporal heterogeneity) and scale dependence
(spatio-temporal heterogeneity; (Vranken et al. 2015)). It is mainly
derived from information theory and thermodynamics, including besides
others Shannon entropy (and its analogous forms), Boltzmann entropy, or
Renyi entropy. However, existing studies show that insights do not only
depend on the formulation of the used entropy, but also on the
underlining data representation, for example the composition of
categories or the co-occurrence matrix (Zhao and Zhang 2019).

The Shannon diversity index (Shannon 1948) quantifies the richness and
evenness of categories in the landscape. Shannon's entropy formula can
also be modified to include the landscape's spatial configuration, e.g.,
by weights calculated from intra- and inter-class distances (Claramunt
2012). However, to quantify the spatial configuration and the total
complexity of the landscape, other measures from information theory can
be adapted. Nowosad and Stepinski (2019) proposed to treat categorical
raster as a bivariate random variable \((x,y)\) where \(x\) is the
category of the focal cell and \(y\) is the category of the neighboring
cell. This approach allows us to compress information about the
landscape's composition and configuration into a co-occurrence matrix,
which could be used to calculate various entropy measures. This includes
conditional entropy (representing configurational complexity), joint
entropy (representing overall spatial-thematic complexity), mutual
information, and relative mutual information (representing the degree of
spatial autocorrelation).

The last few years have seen a surge in the development of entropy-based
metrics for landscape ecology based on notions from thermodynamics
initiated by the work of Vranken et al. (2015) and Cushman (2015). In
the late 1800s, Boltzmann formulated a probabilistic interpretation of
the second law of thermodynamics, using the concepts of macrostate (the
general state of a system) and microstate (the configuration of the
system elements). Cushman (2016) proposed to relate the edge length to
the microstate of the landscape, and use the proportion of microstates
to compute the relative Boltzmann entropy of a landscape mosaic. This
approach was generalized for calculations based on raster's surface
model and point patterns (Cushman 2021). Subsequently, Gao et al. (2017)
proposed to use the Boltzmann entropy to quantify the complexity of a
landscape surface by transforming the input raster into a series of
landscape surfaces with different levels of detail (microstate) and
calculating the Boltzmann entropy based on the total number of
microstates that are able to generate the observed macrostate. Recently,
Zhang et al. (2020) extended the original definition of Boltzmann
entropy to incorporate the information about the adjacency of the same
categories in the landscape mosaic by using the number of contiguous
patches of the same category.

Moreover, many other entropy-based metrics have been proposed for use in
landscape ecology. The Renyi (Rényi 1961) and Gibbs entropies, which are
both generalizations of the Shannon entropy, are applied to quantify the
complexity of the landscape. The Rao quadratic entropy (Rao 1982) has
also been applied recently (Rocchini et al. 2017), as it measures not
only the relative abundances of elements, but also the pairwise
dissimilarities or distances between them. Thus, it can be useful in
cases, for example, when the dissimilarities between land cover types,
and not only their categories, are relevant. Another recent development
is the use of Kullback-Leibler divergence (also known as relative
entropy), which is a measure of differences between two probability
distributions to describe patterns across scales (Huckeba et al. 2023).

\subsection{Landscape mosaic}\label{landscape-mosaic}

\subsection{Vector metrics}\label{vector-metrics}

Many landscape ecology studies are based on raster data. However, the
decision on the data model is often driven by the availability of the
data and software, or the familiarity with the approach (Costanza et al.
2019). For example, several early landscape ecology studies were based
on polygonal vector data that was converted to raster data . Yet,
nowadays, vector data is increasingly available, and most geospatial
software supports it.

Hundreds of metrics have already been developed for raster data, so the
most straightforward approach is to reimplement them for vector data.
This is possible for most of the metrics, with only several exceptions
of metrics that are specifically related to raster data (Yao et al.
2022). Yet, the vector data model has been widely used in other fields,
such as geography or urban studies. For example, in urban planning,
vector metrics are applied to quantify the shapes of urban areas and
characterize the complexity of building footprints (Basaraner and
Cetinkaya 2017). Earlier approaches included four categories of
compactness measures for vector data, namely perimeter-area, single
parameters of related circles, dispersion of elements of the area around
a centroid, and direct comparison to standard shapes (Maceachren 1985).
More recently, a unified theoretical foundation for measuring shape
compactness was introduced using a set of ten distinct properties of a
circle and metrics associated with each of these properties (Angel et
al. 2010)

Vector-based metrics also have their limitations. The most important is
the computational complexity, which makes the calculation of vector
metrics slower than their raster equivalents. Another technical issue is
the need for the data to be topologically correct, which is not always
the case, especially for data from different sources. Last, the
pattern-process link is still less explored for vector-based metrics
compared to raster-based metrics.

\subsection{Operations on spatial
patterns}\label{operations-on-spatial-patterns}

Spatial patterns can be analyzed through a range of computational
operations, such as comparing, searching, or grouping (Turner 1990;
Wickham and Norton 1994; Jasiewicz et al. 2015; Remmel 2020; Nowosad
2021). These operations are based on spatial signatures which are
multi-numerical representations of landscape pattern, and dissimilarity
measures which are functions that quantify differences between the
signatures.

Spatial patterns can be compared between different areas, or for the
same area at two different moments in time. Comparing spatial patterns
is often used to analyze landscape dynamics, e.g., to detect changes in
landscape structure over time (Duncan and Boruff 2023). Furthermore,
pattern-based search is a one-to-many operation in which the spatial
signature of a query area is compared to the signatures of all other
areas. This allows to find areas with similar patterns compared to the
query area, e.g., areas with similar environmental conditions (Jasiewicz
et al. 2014). Finally, the grouping of spatial patterns is a
many-to-many operation in which spatial signatures are calculated for
all areas and clustered based on their similarity (Nowosad and Stepinski
2021).

\section{Simulation models}\label{simulation-models}

Simulation models are a powerful tool to study complex adaptive
ecological systems in controllable, reproducible, and replicable
settings (Pascual 2005; Petrovskii and Petrovskaya 2012). Thus, they can
be seen as experimental systems that allow, unlike to nature, all
imaginable manipulations in order to advance theoretical developments or
test hypotheses (Peck 2004). Due to the usual temporal and spatial
scales, complex interactions and feedbacks, or scale mismatches between
patterns and processes, simulation models are one of the major
approaches in landscape ecology (Schröder and Seppelt 2006; Synes et al.
2016). In general, simulations models can be classified using two major
divisions, namely as \emph{i)} predictive or exploratory models, and
\emph{ii)} pattern- or process-based models (Peck 2004; Synes et al.
2016). Here, we are going to focus mainly on exploratory models, but
include both pattern- and process-based models.

\subsection{Pattern based}\label{pattern-based}

\subsubsection{Neutral landscapes
models}\label{neutral-landscapes-models}

Neutral landscapes models simulate landscape patterns without assuming
any underlying abiotic or biotic processes. Thus, they are often used as
null hypothesis or baselines for comparisons with real landscape
patterns, while controlling certain aspects of the landscape (Li et al.
2004; Wang and Malanson 2008). The first neutral models simply assigned
land cover types randomly to cells in the landscape and were based on
percolation theory. Further developments included hierarchical models
that considered different spatial scales while assigning land cover
types. While these first two classes can only simulate landscape with
discrete land cover classes, fractal neutral models can also simulate
continuous land cover maps by utilizing e.g., Brownian motion. Last,
more complex neutral models were developed that are based on contagion,
autocorrelation, or spatial aggregation concepts to simulate more
realistic landscapes (Wang and Malanson 2008).

doi:10.12942/lrlr-2014-2 Perlin noise (Etherington 2022) Binary space
partitioning {[}Etherington2022a{]} Generating in comparison to real
landscapes https://doi.org/10.1016/j.ecoinf.2017.08.003
https://doi.org/10.1016/j.ecolmodel.2018.12.010

In order to make multiple neutral landscape model algorithms accessible,
several software libraries are available in prominent programming
languages including Python {[}NLMpy; Etherington et al. (2015){]}, R
{[}NLMR; Sciaini et al. (2018){]}, or Julia
(https://github.com/EcoJulia/NeutralLandscapes.jl).

\subsubsection{Cellular Automata}\label{cellular-automata}

\subsection{Process based}\label{process-based}

\href{https://doi.org/10.1111/ecog.05687}{RangeShifter}

\subsubsection{Landscape Generator}\label{landscape-generator}

\href{https://doi.org/10.1371/journal.pone.0064968}{G-RaFFE}
\href{https://doi.org/10.1002/ece3.2145}{LG}
https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0222949\&type=printable

\subsubsection{Coupled lattice}\label{coupled-lattice}

\subsubsection{IBMs/ABMs}\label{ibmsabms}

\subsection{Misc}\label{misc}

\href{https://doi.org/10.1016/j.ecolmodel.2019.108837}{Marine SDMS}
\href{https://doi.org/10.1111/jbi.14617}{SDM past present future}

\href{https://doi.org/10.1111/j.1365-294X.2010.04678.x}{Simulations in
landscape genetics}
\href{https://doi.org/10.1007/s10980-017-0605-9}{HexSim modeling
environment} \href{https://doi.org/10.1016/j.tree.2023.04.010}{Digital
twins?}

\newpage{}

\subsubsection{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-Angel2010}
Angel S, Parent J, Civco DL (2010) Ten compactness properties of
circles: Measuring shape in geography. Canadian Geographies /
G{é}ographies canadiennes 54:441--461.
\url{https://doi.org/10.1111/j.1541-0064.2009.00304.x}

\bibitem[\citeproctext]{ref-Basaraner2017}
Basaraner M, Cetinkaya S (2017) Performance of shape indices and
classification schemes for characterising perceptual shape complexity of
building footprints in GIS. International Journal of Geographical
Information Science 31:1952--1977.
\url{https://doi.org/10.1080/13658816.2017.1346257}

\bibitem[\citeproctext]{ref-Chi2016}
Chi M, Plaza A, Benediktsson JA, et al (2016) Big data for remote
sensing: Challenges and opportunities. Proceedings of the IEEE
104:2207--2219. \url{https://doi.org/10.1109/JPROC.2016.2598228}

\bibitem[\citeproctext]{ref-Claramunt2012}
Claramunt C (2012)
\href{https://doi.org/10.1007/978-3-642-33999-8_28}{Towards a
spatio-temporal form of entropy}. In: Advances in conceptual modeling.
Springer Berlin Heidelberg, Berlin, Heidelberg, pp 221--230

\bibitem[\citeproctext]{ref-Costanza2019}
Costanza JK, Riitters K, Vogt P, Wickham J (2019) Describing and
analyzing landscape patterns: Where are we now, and where are we going?
Landscape Ecology 34:2049--2055.
\url{https://doi.org/10.1007/s10980-019-00889-6}

\bibitem[\citeproctext]{ref-Cushman2015}
Cushman SA (2015) Thermodynamics in landscape ecology: The importance of
integrating measurement and modeling of landscape entropy. Landscape
Ecology 30:7--10. \url{https://doi.org/10.1007/s10980-014-0108-x}

\bibitem[\citeproctext]{ref-Cushman2016}
Cushman SA (2016) Calculating the configurational entropy of a landscape
mosaic. Landscape Ecology 31:481--489.
\url{https://doi.org/10.1007/s10980-015-0305-2}

\bibitem[\citeproctext]{ref-Cushman2021a}
Cushman SA (2021) Generalizing boltzmann configurational entropy to
surfaces, point patterns and landscape mosaics. Entropy 23:1616.
\url{https://doi.org/10.3390/e23121616}

\bibitem[\citeproctext]{ref-Dale1999}
Dale MRT (1999) \href{https://doi.org/10.1017/CBO9780511612589}{Spatial
pattern analysis in plant ecology}, 1st edn. Cambridge University Press

\bibitem[\citeproctext]{ref-Duncan2023}
Duncan JMA, Boruff B (2023) Monitoring spatial patterns of urban
vegetation: A comparison of contemporary high-resolution datasets.
Landscape and Urban Planning 233:104671.
\url{https://doi.org/10.1016/j.landurbplan.2022.104671}

\bibitem[\citeproctext]{ref-Etherington2022}
Etherington TR (2022) Perlin noise as a hierarchical neutral landscape
model. Web Ecology 22:1--6. \url{https://doi.org/10.5194/we-22-1-2022}

\bibitem[\citeproctext]{ref-Etherington2015}
Etherington TR, Holland EP, O'Sullivan D (2015) NLMpy: A python software
package for the creation of neutral landscape models within a general
numerical framework. Methods in Ecology and Evolution 6:164--168.
\url{https://doi.org/10.1111/2041-210X.12308}

\bibitem[\citeproctext]{ref-Forman1995}
Forman RTT (1995) Land mosaics: The ecology of landscapes and regions.
Cambridge University Press, Cambridge, UK

\bibitem[\citeproctext]{ref-Forman1986}
Forman RTT, Godron M (1986) Landscape ecology. Wiley \& Sons,
Chichester, UK

\bibitem[\citeproctext]{ref-Gao2017}
Gao P, Zhang H, Li Z (2017) A hierarchy-based solution to calculate the
configurational entropy of landscape gradients. Landscape Ecology.
\url{https://doi.org/10.1007/s10980-017-0515-x}

\bibitem[\citeproctext]{ref-Gergel2017}
Gergel SE, Turner MG (eds) (2017)
\href{https://doi.org/10.1007/978-1-4939-6374-4}{Learning landscape
ecology}. Springer New York, New York, NY

\bibitem[\citeproctext]{ref-Grimm1996}
Grimm V, Frank K, Jeltsch F, et al (1996) Pattern-oriented modelling in
population ecology. The Science of the Total Environment 183:151--166.
\url{https://doi.org/10.1016/0048-9697(95)04966-5}

\bibitem[\citeproctext]{ref-Huckeba2023}
Huckeba G, Andresen B, Roach TNF (2023) Multi-scale spatial ecology
analyses: A kullback information approach. Landscape Ecology
38:645--657. \url{https://doi.org/10.1007/s10980-022-01514-9}

\bibitem[\citeproctext]{ref-Jaric2020}
Jarić I, Correia RA, Brook BW, et al (2020) iEcology: Harnessing large
online resources to generate ecological insights. Trends in Ecology \&
Evolution 35:630--639. \url{https://doi.org/10.1016/j.tree.2020.03.003}

\bibitem[\citeproctext]{ref-Jasiewicz2015}
Jasiewicz J, Netzel P, Stepinski T (2015) GeoPAT: A toolbox for
pattern-based information retrieval from large geospatial databases.
Computers \& Geosciences 80:62--73.
\url{https://doi.org/10.1016/j.cageo.2015.04.002}

\bibitem[\citeproctext]{ref-Jasiewicz2014}
Jasiewicz J, Netzel P, Stepinski TF (2014) Landscape similarity,
retrieval, and machine mapping of physiographic units. Geomorphology
221:104--112. \url{https://doi.org/10.1016/j.geomorph.2014.06.011}

\bibitem[\citeproctext]{ref-Li2004}
Li X, He HS, Wang X, et al (2004) Evaluating the effectiveness of
neutral landscape models to represent a real landscape. Landscape and
Urban Planning 69:137--148.
\url{https://doi.org/10.1016/j.landurbplan.2003.10.037}

\bibitem[\citeproctext]{ref-Maceachren1985}
Maceachren AM (1985) Compactness of geographic shape: Comparison and
evaluation of measures. Geografiska Annaler: Series B, Human Geography
67:53--67. \url{https://doi.org/10.1080/04353684.1985.11879515}

\bibitem[\citeproctext]{ref-Nathan2022}
Nathan R, Monk CT, Arlinghaus R, et al (2022) Big-data approaches lead
to an increased understanding of the ecology of animal movement. Science
375:eabg1780. \url{https://doi.org/10.1126/science.abg1780}

\bibitem[\citeproctext]{ref-Nedd2021}
Nedd R, Light K, Owens M, et al (2021) A synthesis of land use/land
cover studies: Definitions, classification systems, meta-studies,
challenges and knowledge gaps on a global landscape. Land 10:994.
\url{https://doi.org/10.3390/land10090994}

\bibitem[\citeproctext]{ref-Newman2019}
Newman EA, Kennedy MC, Falk DA, McKenzie D (2019) Scaling and complexity
in landscape ecology. Frontiers in Ecology and Evolution 7:293.
\url{https://doi.org/10.3389/fevo.2019.00293}

\bibitem[\citeproctext]{ref-Nowosad2021a}
Nowosad J (2021) Motif: An open-source r tool for pattern-based spatial
analysis. Landscape Ecology 36:29--43.
\url{https://doi.org/10.1007/s10980-020-01135-0}

\bibitem[\citeproctext]{ref-Nowosad2019a}
Nowosad J, Stepinski TF (2019) Information theory as a consistent
framework for quantification and classification of landscape patterns.
Landscape Ecology 34:2091--2101.
\url{https://doi.org/10.1007/s10980-019-00830-x}

\bibitem[\citeproctext]{ref-Nowosad2021b}
Nowosad J, Stepinski TF (2021) Pattern-based identification and mapping
of landscape types using multi-thematic data. International Journal of
Geographical Information Science 35:1634--1649.
\url{https://doi.org/10.1080/13658816.2021.1893324}

\bibitem[\citeproctext]{ref-Pascual2005}
Pascual M (2005) Computational ecology: From the complex to the simple
and back. PLoS Computional Biology 1:e18.
\url{https://doi.org/10.1371/journal.pcbi.0010018}

\bibitem[\citeproctext]{ref-Peck2004}
Peck SL (2004) Simulation as experiment: A philosophical reassessment
for biological modeling. Trends in Ecology \& Evolution 19:530--534.
\url{https://doi.org/10.1016/j.tree.2004.07.019}

\bibitem[\citeproctext]{ref-Petrovskii2012}
Petrovskii S, Petrovskaya N (2012) Computational ecology as an emerging
science. Interface Focus 2:241--254.
\url{https://doi.org/10.1098/rsfs.2011.0083}

\bibitem[\citeproctext]{ref-Poisot2019}
Poisot T, LaBrie R, Larson E, et al (2019) Data-based, synthesis-driven:
Setting the agenda for computational ecology. Ideas in Ecology and
Evolution 12: \url{https://doi.org/10.24908/iee.2019.12.2.e}

\bibitem[\citeproctext]{ref-Prlic2012}
Prlić A, Procter JB (2012) Ten simple rules for the open development of
scientific software. PLoS Computational Biology 8:e1002802.
\url{https://doi.org/10.1371/journal.pcbi.1002802}

\bibitem[\citeproctext]{ref-Rao1982}
Rao CR (1982) Diversity and dissimilarity coefficients: A unified
approach. Theoretical Population Biology 21:24--43.
\url{https://doi.org/10.1016/0040-5809(82)90004-1}

\bibitem[\citeproctext]{ref-Remmel2020}
Remmel TK (2020) Distributions of hyper-local configuration elements to
characterize, compare, and assess landscape-level spatial patterns.
Entropy 22:420. \url{https://doi.org/10.3390/e22040420}

\bibitem[\citeproctext]{ref-Renyi1961}
Rényi A (1961) On measures of entropy and information. In: Proceedings
of the fourth berkeley symposium on mathematical statistics and
probability, volume 1: Contributions to the theory of statistics.
University of California Press, pp 547--562

\bibitem[\citeproctext]{ref-Risser1984}
Risser PG, Karr JR, Forman RTT (1984) Landscape ecology: Directions and
approaches. Illinois Natural History Survery Special Publication 2:7--14

\bibitem[\citeproctext]{ref-Rocchini2017}
Rocchini D, Marcantonio M, Ricotta C (2017) Measuring rao's q diversity
index from remote sensing: An open source solution. Ecological
Indicators 72:234--238.
\url{https://doi.org/10.1016/j.ecolind.2016.07.039}

\bibitem[\citeproctext]{ref-Schroder2006}
Schröder B, Seppelt R (2006) Analysis of pattern--process interactions
based on landscape models - overview, general concepts, and
methodological issues. Ecological Modelling 199:505--516.
\url{https://doi.org/10.1016/j.ecolmodel.2006.05.036}

\bibitem[\citeproctext]{ref-Sciaini2018}
Sciaini M, Fritsch M, Scherer C, Simpkins CE (2018) NLMR and
landscapetools: An integrated environment for simulating and modifying
neutral landscape models in r. Methods in Ecology and Evolution
9:2240--2248. \url{https://doi.org/10.1111/2041-210X.13076}

\bibitem[\citeproctext]{ref-Shannon1948}
Shannon CE (1948) A mathematical theory of communication. Bell System
Technical Journal 27:379--423.
\url{https://doi.org/10.1002/j.1538-7305.1948.tb01338.x}

\bibitem[\citeproctext]{ref-Synes2016}
Synes NW, Brown C, Watts K, et al (2016) Emerging opportunities for
landscape ecological modelling. Current Landscape Ecology Reports
1:146--167

\bibitem[\citeproctext]{ref-Turner1989}
Turner MG (1989) Landscape ecology: The effect of pattern on process.
Annual Review of Ecology and Systematics 20:171--197.
\url{https://doi.org/10.1146/annurev.es.20.110189.001131}

\bibitem[\citeproctext]{ref-Turner1990}
Turner MG (1990) Spatial and temporal analysis of landscape patterns.
Landscape Ecology 4:21--30. \url{https://doi.org/10.1007/BF02573948}

\bibitem[\citeproctext]{ref-Turner2005}
Turner MG (2005) Landscape ecology: What is the state of the science?
Annual Review of Ecology, Evolution, and Systematics 36:319--344.
\url{https://doi.org/10.1146/annurev.ecolsys.36.102003.152614}

\bibitem[\citeproctext]{ref-Turner2015}
Turner MG, Gardner RH (2015) Landscape ecology in theory and practice:
Pattern and process, 2nd edition. Springer, New York

\bibitem[\citeproctext]{ref-vonKrogh2006}
von Krogh G, von Hippel E (2006) The promise of research on open source
software. Management Science 52:975--983.
\url{https://doi.org/10.1287/mnsc.1060.0560}

\bibitem[\citeproctext]{ref-Vranken2015}
Vranken I, Baudry J, Aubinet M, et al (2015) A review on the use of
entropy in landscape ecology: Heterogeneity, unpredictability, scale
dependence and their links with thermodynamics. Landscape Ecology
30:51--65. \url{https://doi.org/10.1007/s10980-014-0105-0}

\bibitem[\citeproctext]{ref-Wang2008}
Wang Q, Malanson GP (2008) Neutral landscapes: Bases for exploration in
landscape ecology. Geography Compass 2:319--339.
\url{https://doi.org/10.1111/j.1749-8198.2008.00090.x}

\bibitem[\citeproctext]{ref-Wickham1994}
Wickham JD, Norton DJ (1994) Mapping and analyzing landscape patterns.
Landscape Ecology 9:7--23. \url{https://doi.org/10.1007/BF00135075}

\bibitem[\citeproctext]{ref-Wiegand2003}
Wiegand T, Jeltsch F, Hanski I, Grimm V (2003) Using pattern-oriented
modeling for revealing hidden information: A key for reconciling
ecological theory and application. Oikos 100:209--222.
\url{https://doi.org/10.1034/j.1600-0706.2003.12027.x}

\bibitem[\citeproctext]{ref-Wiens1997}
Wiens JA (1997)
\href{https://doi.org/10.1016/B978-012323445-2/50005-5}{Metapopulation
dynamics and landscape ecology}. In: Metapopulation biology. Elsevier,
pp 43--62

\bibitem[\citeproctext]{ref-Wiersma2022}
Wiersma YF (2022) A review of landscape ecology experiments to
understand ecological processes. Ecological Processes 11:57.
\url{https://doi.org/10.1186/s13717-022-00401-0}

\bibitem[\citeproctext]{ref-Wilson2014}
Wilson G, Aruliah DA, Brown CT, et al (2014) Best practices for
scientific computing. PLoS Biology 12:e1001745.
\url{https://doi.org/10.1371/journal.pbio.1001745}

\bibitem[\citeproctext]{ref-With2019}
With KA (2019) Essentials of landscape ecology, 1st edn. Oxford
University Press, Oxford, UK

\bibitem[\citeproctext]{ref-Wu2013}
Wu J (2013) Key concepts and research topics in landscape ecology
revisited: 30 years after the allerton park workshop. Landscape Ecology
28:1--11. \url{https://doi.org/10.1007/s10980-012-9836-y}

\bibitem[\citeproctext]{ref-Wu2002}
Wu J, Hobbs R (2002) Key issues and research priorities in landscape
ecology: An idiosyncratic synthesis. Landscape Ecology 17:355--365.
\url{https://doi.org/10.1023/A:1020561630963}

\bibitem[\citeproctext]{ref-Wuest2020}
Wüest RO, Zimmermann NE, Zurell D, et al (2020) Macroecology in the age
of big data -- where to go from here? Journal of Biogeography 47:1--12.
\url{https://doi.org/10.1111/jbi.13633}

\bibitem[\citeproctext]{ref-Yao2022}
Yao Y, Cheng T, Sun Z, et al (2022) VecLI: A framework for calculating
vector landscape indices considering landscape fragmentation.
Environmental Modelling and Software 149:105325.
\url{https://doi.org/10.1016/j.envsoft.2022.105325}

\bibitem[\citeproctext]{ref-Zhang2020}
Zhang H, Wu Z, Lan T, et al (2020) Calculating the wasserstein
metric-based boltzmann entropy of a landscape mosaic. Entropy 22:381.
\url{https://doi.org/10.3390/e22040381}

\bibitem[\citeproctext]{ref-Zhao2019}
Zhao Y, Zhang X (2019) Calculating spatial configurational entropy of a
landscape mosaic based on the wasserstein metric. Landscape Ecology
34:1849--1858. \url{https://doi.org/10.1007/s10980-019-00876-x}

\end{CSLReferences}

\end{document}