### 2024-06-24 Sections/simulations.qmd
Update simulations

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth, Schouten -->

# Simulation models

Simulation models are a powerful tool to study complex adaptive ecological systems in controllable, reproducible, and replicable settings [@Pascual2005; @Petrovskii2012].
Thus, they can be seen as experimental systems that allow, unlike to the natural world, all imaginable manipulations in order to advance theoretical developments or test hypotheses [@Peck2004].
Due to the usual temporal and spatial scales, complex interactions and feedbacks, or scale mismatches between patterns and processes, simulation models are one of the major approaches in landscape ecology [@Schroder2006; @Synes2016].
In general, simulations models can be classified using two major divisions, namely as *i)* predictive or exploratory models, and *ii)* pattern- or process-based models [@Peck2004;@Synes2016].
Here, we are going to focus mainly on exploratory models, but include both pattern- and process-based models.

## Neutral landscapes models
Neutral landscapes models simulate landscape patterns without assuming any underlying abiotic or biotic processes.
Thus, they are often used as null hypothesis, baselines, or scenarios for comparisons with real landscape patterns, while controlling certain aspects of the landscape [@Li2004; @Wang2008].
The first neutral models simply assigned land cover types randomly to cells in the landscape and were based on percolation theory [@Gardner1987].
Further developments included hierarchical models that considered different spatial scales while assigning land cover types [@ONeill1992].
While these first two classes can only simulate landscape with discrete land cover classes, fractal neutral models can also simulate continuous land cover maps by utilizing e.g., Brownian motion [@Palmer1992].
Borrowing from computer graphics, more recent neutral landscape models make use of spectral synthesis [e.g., Perlin noise\; @Etherington2022] or Binary space partitioning [@Etherington2022a] hierarchical neutral landscapes.
In order to simulate landscapes dominated by human activities approaches based on least-cost network movement exist [@Etherington2024].
Neutral landscapes models can also be used to simulated landscapes similar to real landscapes [@Inkoom2017], or based on target values of landscape metrics [@VanStrien2016; @Justeau-Allaire2022].
<!-- In order to make multiple neutral landscape model algorithms accessible, several software libraries are available in prominent programming languages including Python [NLMpy\; @Etherington2015], R [NLMR\; @Sciaini2018], or Julia (https://github.com/EcoJulia/NeutralLandscapes.jl). -->

## Landscape Generator
Processes-based landscape generators are related to neutral landscapes models because they can simulate virtual landscapes.
However, in contrast to neutral landscape models, processes-based landscape generators are able to simulate structurally more realistic landscapes by including pattern creating processes [@Langhammer2019; @Salecker2019].

<!--
Cellular automaton
[G-RaFFE](https://doi.org/10.1371/journal.pone.0064968)
[GradientLand Software](https://doi.org/10.1016/j.ecoinf.2014.12.001)
[PioLaG](https://doi.org/10.1007/s10980-020-01066-w)
-->

## IBMs /Coupled lattice

## Digital twins

<!-- 
[RangeShifter](https://doi.org/10.1111/ecog.05687)
[Marine SDMS](https://doi.org/10.1016/j.ecolmodel.2019.108837)
[SDM past present future](https://doi.org/10.1111/jbi.14617)
[Simulations in landscape genetics](https://doi.org/10.1111/j.1365-294X.2010.04678.x)
[HexSim modeling environment](https://doi.org/10.1007/s10980-017-0605-9)
[Digital twins?](https://doi.org/10.1016/j.tree.2023.04.010)
-->