### 2024-07-11 Sections/08-software.qmd
Update MS and software

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth, Nowosad, Bosch, Jung, Schouten  -->

# Software and programming languages

<!-- "The future applications of landscape pattern metrics will almost certainly depend on the software and analysis tools available" -- @costanzaDescribingAnalyzingLandscape2019 -->

"Open science" involves full transparency and data access at all stages of the research processes and the integrative and multi-disciplinary field of ecology can massively benefit from this concept [@Hampton2015].
One key aspect of achieving open science is to use open source software, which ensures transparency and consequently increasing the (public) trust into the analyses [@Powers2019].
Related to this, script-based workflows have the advantage over workflows involving graphical user interfaces that the former are shareable and reproducible [@Hampton2015; @Powers2019].
Thus, we will focus the following section on script-based open-source software suitable for landscape ecology.
Given the increasing availability of data and complexity of methods, future progress within the field of landscape ecology will depend on the software and analysis tools available [@Costanza2019]

<!-- R: Hesselbarth -->
The open-source and script-based R programming languages [@RCoreTeam2024] was first released in 1995 [@Smith2016] with a focus on statistical computing and data visualization.
However, due to its extensions (also called packages), the language is popular in many academic fields including ecology [@Lai2019] and it is suitable for spatial data analysis or landscape ecology [e.g., @Bivand2020; @Hesselbarth2021].

<!-- 
landscapemetrics
geodiv
vectormetrics (also: VecLI -- they mention open source in their paper a few times, but the software seems to be closed source -- https://urbancomp.net/archives/vecliv3beta)
bespatial
LandComp
motif
?GLCMTextures
?rasterdiv
?terra 
-->

<!-- Python: Bosch -->

<!--Julia: Schouten -->

<!--Others: -->

<!-- 
GuidosToolbox
?GRASS GIS
?QGIS: LecoS plugin 
-->