### 2024-07-12 Sections/04-sampling.qmd
Moving potential literature to github issue

---
bibliography: CLER.bib
---

<!-- Person responsible: Bosch -->

# Sampling in landscape ecology