### 2024-07-12 Sections/05-simulations.qmd
Update digital twins section

---
bibliography: CLER.bib
---

<!-- Person responsible: Hesselbarth, Schouten -->

# Simulation models

Simulation models are a powerful tool to study complex adaptive ecological systems in controllable, reproducible, and replicable settings [@Pascual2005; @Petrovskii2012].
Thus, they can be seen as experimental systems that allow, unlike to the natural world, all imaginable manipulations in order to advance theoretical developments or test hypotheses [@Peck2004].
Due to the usual temporal and spatial scales, complex interactions and feedbacks, or scale mismatches between patterns and processes, simulation models are one of the major approaches in landscape ecology [@Schroder2006; @Synes2016].
In general, simulations models can be classified using two major divisions, namely as *i)* predictive or exploratory models, and *ii)* pattern- or process-based models [@Peck2004;@Synes2016].
Here, we are going to focus mainly on exploratory models, but include both pattern- and process-based models.

## Landscapes simulators

Landscape simulators are typically used as null hypotheses, baselines, or scenarios because they allow to control certain aspects of the landscape [@Li2004; @Wang2008] and are generally simpler than ecological simulations models [@Hess2020].
Landscape simulators can be classified into two major categories, namely pattern-based and process-based approaches [@Peer2013; Langhammer2019].

Pattern-based approaches simulate landscapes without assuming any underlying abiotic or biotic processes [i.e., neutral landscape model\; @Gardner1987].
The earliest neutral models are based on percolation theory and assign LULC types randomly to cells in the landscape[@Gardner1987] or hierarchical models that consider different spatial scales while assigning cell values [@ONeill1992].
Landscapes characterized by continuous values can be simulated by fractal models, such as Brownian motion [@Palmer1992].
Borrowing from computer graphics, more recent neutral landscape models make use of spectral synthesis [e.g., Perlin noise\; @Etherington2022] or binary space partitioning [@Etherington2022a].
Neutral landscape models are also able to simulate landscapes dominated by anthropogenic activities using based on least-cost networks [@Etherington2024].
In order to ensure realistic neutral landscapes several approaches exist that use comparisons with real landscapes [@Inkoom2017] or are based on target values optimization [@VanStrien2016; @Justeau-Allaire2022].
<!-- In order to make multiple neutral landscape model algorithms accessible, several software libraries are available in prominent programming languages including Python [NLMpy\; @Etherington2015], R [NLMR\; @Sciaini2018], Julia (https://github.com/EcoJulia/NeutralLandscapes.jl), or GUI software (GradientLand, QRULE) -->

Contrastingly, processes-based approaches explicitly include abiotic or biotic pattern-forming processes [i.e., landscape generators\; @Langhammer2019; @Salecker2019].
Several earlier landscape generators are based on cellular automata models and are able to simulate urban growth [@Clarke1998] or deforestation [@Soares-Filho2002].
More recent landscape generators allow to simulate patchy landscapes based on Gibbs processes [@Gaucherel2006], deforestation based on road and agricultural fields access [@Peer2013], agricultural areas dominated by smallholders [@Salecker2019], or vegetation surround watering points in semi-arid savanna rangelands [combining pattern- and process-bases approaches\; @Hess2020].

## Individual-based models

Individual-based models (or agent-based models) simulate discrete entities that are described by attributes and behaviors and patterns emerge from bottom-up interactions of individuals with each other and their environment [@DeAngelis2014].
In landscape ecology individual-based models are increasingly used to model especially social-ecological systems [e.g, @Carauta2018; @Dislich2018; @Gonzalez-Redin2019], but also disturbances [e.g., @Seidl2017a; @Spies2017; @Kamimura2019] or connectivity [e.g., @Day2020].
In order to facilitate future model developments, recent progress includes adaptable modeling frameworks [@Schumaker2018] or reusable building block [@Berger2024].
Furthermore, hybrid approaches that couple or even integrated different model types with individual-based models could benefit from rich development history or facilitate corresponding strengths [@OSullivan2016], e.g., linking an individual-based and ecosystem model [@Polhill2011].

## Digital twins

Digital Twins include three connected elements, namely a physical object or system, its digital representation, and a data exchange between the former two [@Jones2020; @DeKoning2023] and there is an increasing interest and use of them in both the industry and academia [@Jones2020].
The need for timely and evidence-based decision making in combination with increasing data availability makes them a powerful tool for ecology [@DeKoning2023].
Digital twins are increasingly applied in agricultural landscapes research, including topics such as crop and land cultivation, livestock farming, controlled environment farming, or fertilization management [@Cesco2023; @Purcell2023].
Further applications include planning of rural ecological landscapes [@Tan2024], or exploration of relationships between urban expansion and vegetation coverage [@Zhao2022].
The use of general data processing, statistical and mechanistic models, or artificial intelligence allows them to constantly update the digital representation and analyze, visualize, or predict the physical counterpart.
Thus, Digital Twins can provide real-time state of nature measurements, early detection of conservation trends, relationships of ecosystem trends and environmental conditions, impact assessments of interventions, or identify uncertainties and information gaps [@DeKoning2023].