### 2024-08-29 Sections/07-spatial-planning.qmd
Updating software section and tiny cosmetic changes

---
bibliography: CLER.bib
---

<!-- Person responsible: Jung -->

# Spatial planning