### 2024-08-29 Manuscript.tex
Update Bib and manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paperpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[margin=25mm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{lineno}\linenumbers
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Computational methods in landscape ecology},
  pdfauthor={Maximilian H.K. Hesselbarth; Jakub Nowosad; Marti Bosch; Martin Jung; Rafael Schouten},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Computational methods in landscape ecology}

  \author{Maximilian H.K. Hesselbarth}
            \affil{%
                  International Institute for Applied Systems Analysis,
                  Biodiversity, Ecology, and Conservation Group,
                  Laxenburg, Austria
              }
        \author{Jakub Nowosad}
            \affil{%
                  Adam Mickiewicz University, Poznan, Poland
              }
          \affil{%
                  University of Münster, Münster, Germany
              }
        \author{Marti Bosch}
            \affil{%
                  École polytechnique fédérale de Lausanne, Lausanne,
                  Switzerland
              }
        \author{Martin Jung}
            \affil{%
                  International Institute for Applied Systems Analysis,
                  Biodiversity, Ecology, and Conservation Group,
                  Laxenburg, Austria
              }
        \author{Rafael Schouten}
            \affil{%
                  Globe Institute, University of Copenhagen, Copenhagen,
                  Denmark
              }

\date{}
\begin{document}
\maketitle

\setstretch{1.5}
\section{Introduction}\label{introduction}

Landscapes are typically defined as mosaics of different land covers,
habitats, ecosystems, or land use systems (Forman and Godron 1986;
Forman 1995) - with empathize on heterogeneity of at least one factor of
interest (Turner and Gardner 2015). Linking spatial heterogeneity and
ecological processes, including potential feedbacks, is the fundamental
concept of landscape ecology (Risser et al. 1984; Turner 1989, 2005).
Typical research topics include ecological flows in landscape mosaics,
drivers and consequences of land use and land cover (LULC) change,
nonlinear dynamics and landscape complexity, scale and scaling, spatial
analysis and landscape modeling, linking landscape metrics to ecological
processes, integrating humans and their activities into landscape
ecology, landscape conservation and sustainability, or accuracy
assessment and uncertainty analysis (Wu and Hobbs 2002; Wu 2013).

Computational science involves analyzing abstracted core mechanisms of
research questions using data and algorithms. Thus, computational
ecology can be defined as computational science that is used to address
ecological research questions with focus on complex, adaptive systems
and data-driven approaches (Poisot et al. 2019). Computational ecology
in general and simulations specifically are of importance for ecology
because ecological data is often context- and scale-dependent and
unfeasible to study in controllable, reproducible, and replicable
experimental settings (Petrovskii and Petrovskaya 2012; but see Wiersma
2022). Because landscape ecology is a cross-disciplinary field including
(e.g., social sciences, geography, or ecology and evolution; Wiens 1997;
With 2019), the general increase of data availability (Chi et al. 2016;
Jarić et al. 2020; Wüest et al. 2020; Nathan et al. 2022), or the
studied complex landscape systems (Newman et al. 2019), there is the
need for sophisticated computational methods in order to link
heterogeneity patterns and ecological processes. Thus, computational
methods are one of the most important tools of modern scientific
research (Prlić and Procter 2012; Wilson et al. 2014).

For all computational methods the natural world must be translated into
representative data models. The spatial vector data model is based on
points, lines, and polygons features that are defined by their geometry
and additional attributes related to them. Contrastingly, the raster
data model is based on evenly-spaced grid cells and discrete or
continuous values related to each cell (Lovelace et al. 2019; With
2019). The choice of the data models, spatial and temporal scales, or
the thematic resolution are crucial for the analysis (Wade et al. 2003;
Lechner and Rhodes 2016; Nedd et al. 2021).

Here, we aim to introduce the latest developments of computational
methods in landscape ecology. However, we do not provide a systematic
literature review, or a general introduction to (computational)
landscape ecology. For a more comprehensive, general introductions to
landscape ecology, please see Turner and Gardner (2015), Gergel and
Turner (2017), or With (2019). First, we will introduce recent
computational developments of the most important topics of landscape
ecology. This includes, for example, spatial patterns quantification,
connectivity and movement, or model simulations. Second, we will
introduce common software tools that implement or are potentially
capable of implementing these developments. However, due to its many
advantages (Powers and Hampton 2019), we will focus on open-source
software , and specifically scripting languages, i.e., R, Python, and
Julia. Last, we will discuss potential future needs developments of
computational methods in landscape ecology.

\section{Spatial patterns}\label{spatial-patterns}

Spatial patterns can be defined as the scale-dependent predictability of
the physical arrangement of observations (Dale 1999), or as clearly
identifiable structures in nature itself or data extracted from nature
(Grimm et al. 1996). Importantly, patterns observed in nature contain
information about the history of the system, such as demographic
processes, dispersal characteristics, or climatic patterns (Wiegand et
al. 2003). However, spatial patterns are not only a result of processes,
but could also be a driver of them. Thus, untangling the history of a
landscape and linking its spatial patterns to ecological processes is
one of the core concepts of landscape ecology (Turner 1989).

\subsection{Landscape metrics}\label{landscape-metrics}

Traditionally and still the most prominent approaches to quantify
spatial patterns revolve around raster data using categorical categories
based on the patch mosaic model (Frazier and Kedron 2017; Costanza et
al. 2019). The strength of landscape metrics is that they are comparable
easy to apply and communicate (Lausch et al. 2015), however, they are
also critiqued for not being able to realistically represent continuous
features of the natural word (Frazier and Kedron 2017). Metrics can be
calculated on patch-level (describing each patch), class-level
(describing all patches of the same class), or landscape level
(describing the entire landscape) and are able to quantify
characteristics related to e.g., patch areas and edge lengths, patch
shapes, or class aggregation and diversity (Hesselbarth et al. 2019).
Recently also several metrics describing landscape complexity were
introduced, e.g., several entropy measures (see following sections).

\subsection{Entropy}\label{entropy}

It is mainly derived from information theory and thermodynamics, such as
Shannon entropy (and its analogous forms), Boltzmann entropy, or Renyi
entropy. Entropy in landscape ecology is mainly used to quantify the
complexity of the landscape (spatial heterogeneity), and less often
unpredictability (temporal heterogeneity) and scale dependence
(spatio-temporal heterogeneity; Vranken et al. 2015). However, studies
show that insights do not only depend on the formulation of the used
entropy, but also on the underlining data model, for example the
composition of categories or the co-occurrence matrix (Zhao and Zhang
2019).

The Shannon diversity index (Shannon 1948) quantifies the richness and
evenness of categories in the landscape omitting spatial configuration.
The index is based on information theory and is an adaptation of the
Shannon entropy which measures uncertainty of a random variable.
Shannon's entropy formula can also be modified to include the
landscape's spatial configuration, e.g., by weights calculated from
intra- and inter-class distances (Claramunt 2012). However, to quantify
the spatial configuration and the total complexity of the landscape,
other measures from information theory can be adapted. Nowosad and
Stepinski (2019) proposed to treat categorical raster as a bivariate
random variable \((x,y)\) where \(x\) is the category of the focal cell
and \(y\) is the category of the neighboring cell. This approach allows
us to compress information about the landscape's composition and
configuration into a co-occurrence matrix, which could be used to
calculate various entropy measures. This includes conditional entropy
(representing configurational complexity), joint entropy (representing
overall spatial-thematic complexity), mutual information, and relative
mutual information (representing the degree of spatial autocorrelation).

Based on notions from thermodynamics initiated by the work of Cushman
(2015) or Vranken et al. (2015), there was a surge in the development of
entropy-based metrics for landscape ecology in the last few years. In
the late 1800s, Boltzmann formulated a probabilistic interpretation of
the second law of thermodynamics using the concepts of macrostate (the
general state of a system) and microstate (the configuration of the
system elements; Gao and Li 2019). Cushman (2016) proposed to relate the
edge length to the microstate of the landscape, and use the proportion
of microstates to compute the relative Boltzmann entropy of a landscape
mosaic. This approach was generalized for calculations based on raster's
surface model and point patterns (Cushman 2021). Subsequently, Gao et
al. (2017) proposed to use the Boltzmann entropy to quantify the
complexity of a landscape surface by transforming the input raster into
a series of landscape surfaces with different levels of detail
(microstate) and calculating the Boltzmann entropy based on the total
number of microstates that are able to generate the observed macrostate.
Recently, Zhang et al. (2020) extended the original definition of
Boltzmann entropy to incorporate information about the adjacency of the
same categories in the landscape mosaic by using the number of
contiguous patches of the same category.

Moreover, many other entropy-based metrics have been proposed for use in
landscape ecology. The Renyi (Rényi 1961) and Gibbs entropies , which
are both generalizations of the Shannon entropy, are applied to quantify
the complexity of the landscape. The Rao quadratic entropy (Rao 1982)
has also been applied recently (Rocchini et al. 2017), as it measures
not only the relative abundances of elements, but also the pairwise
dissimilarities or distances between them. Thus, it can be useful in
cases, when the dissimilarities between LULC types, and not only their
categories, are relevant. Another recent development is the use of
Kullback-Leibler divergence (also known as relative entropy), which is a
measure of differences between two probability distributions to describe
patterns across scales (Huckeba et al. 2023).

\subsection{Operations on spatial
patterns}\label{operations-on-spatial-patterns}

Spatial patterns can be analyzed through a range of computational
operations, such as comparing, searching, or grouping (Turner 1990;
Wickham and Norton 1994; Jasiewicz et al. 2015; Remmel 2020; Nowosad
2021). These operations are based on spatial signatures which are
multi-numerical representations of landscape pattern, and dissimilarity
measures which are functions that quantify differences between the
signatures.

Spatial patterns can be compared between different areas, or for the
same area at two different moments in time. Comparing spatial patterns
is often used to analyze landscape dynamics, e.g., to detect changes in
landscape structure over time (Duncan and Boruff 2023). Furthermore,
pattern-based search is a one-to-many operation in which the spatial
signature of a query area is compared to the signatures of all other
areas. This allows to find areas with similar patterns compared to the
query area, e.g., areas with similar environmental conditions (Jasiewicz
et al. 2014). Finally, the grouping of spatial patterns is a
many-to-many operation in which spatial signatures are calculated for
all areas and clustered based on their similarity (Nowosad and Stepinski
2021).

\subsection{Landscape mosaic}\label{landscape-mosaic}

Despite the extensive use of the landscape mosaic model and metrics
based on it, major limitations have been identified regarding both its
ability to quantify the spatial structure of landscapes and their
relationship to ecological processes (Gustafson 1998; Li and Wu 2004;
Dramstad 2009; Kupfer 2012; Lausch et al. 2015; Frazier and Kedron
2017). Here, we review four major issues of landscape metrics, namely
their sensitivity to both spatial scale and thematic resolution,
redundancy and interpretation (Dramstad 2009).

\subsubsection{Sensitivity to spatial scale and thematic
resolution}\label{sensitivity-to-spatial-scale-and-thematic-resolution}

Identifying scales at which the ecological processes of interest operate
is a critical first step towards establishing relationships between
pattern and process (Levin 1992; Wu 2007; Cushman and McGarigal 2008).
To that end, a key requirement is to understand the sensitivity of
landscape metrics to spatial scale in terms of both grain and extent.

A key concept to consider regarding grain size is the minimum mapping
unit (MMU), namely the smallest area entity that can be mapped and
analyzed as a discrete patch. Since the input maps should represent not
only the location of patches but also their shape, the grain size should
be several times smaller than the smallest patch (i.e., the MMU) in
order to quantify shapes. Even though the MMU is often assumed as a
natural characteristic of the dataset, it has an impact on the computed
landscape metrics, especially on the number of patches and their size
distribution (Saura 2002).

While the choice of grain is often constrained by the available data,
e.g., Landsat scenes at 30 m resolution, the selection of spatial extent
varies largely depending on the phenomenon of interest, e.g., from large
transects of forests in natural parks to small buffers along the
urban-rural gradient. The spatial extent and sampling strategy is often
subject to many arbitrary criteria that can largely influence results.
This issue has long been recognized in physical geography as the
modifiable areal unit problem {[}MAUP; Openshaw (1979); Openshaw
(1984){]}. Dealing with the MAUP in the context of landscape metrics
include the identification of individual entities that are ecologically
meaningful and not modifiable, the evaluation of ``optimal'' zoning
systems that maximize interzonal variations and minimize intrazonal
variation, and the analysis of the sensitivity of the zoning
configuration (Jelinski and Wu 1996; Duque et al. 2018; Fletcher Jr et
al. 2023).

The scale effects of landscape metrics can be assessed by computing
scalograms, i.e., the response curves of metrics to changing scales.
Some metrics related to the number of patches, edges show consistent
linear and power-law scaling relationships to changing grain, while
other metrics exhibit less predictable responses (Turner et al. 1989;
Saura and Martínez-Millán 2001; Wu et al. 2002; Wu 2004; Saura and
Castro 2007; Šı́mová and Gdulová 2012; Argañaraz and Entraigas 2014;
Frazier et al. 2021). Nonetheless, the effects of changing the sampling
extent can be less predictable than those of changing grain size (Wu
2004). For instance, in landscapes with small extent to grain ratio,
boundary effects are likely to dominate and truncate patches resulting
in unreliable metric computations. In order to address scale issues,
studies should explicitly consider the concept of scope, namely the
ratio of extent to resolution (Schneider 2001). This involves three
fundamental steps, i) considering scope guidelines when designing
experiments, ii) explicitly reporting grain and extent allowing
quantitative scope comparisons with other studies, iii) and comparisons
and replications based on scope (Frazier 2023).

\subsubsection{Redundancy}\label{redundancy}

TODO

\begin{itemize}
\item
  Factor analysis of Riitters et al. (Riitters et al. 1995) found six
  main components: average patch compaction, image texture, average
  patch shape, patch perimeter-area scaling, number of attribute classes
  and large-patch density-area scaling.
\item
  Cushman et al. (Cushman et al. 2008) not only isolate the components
  but also assess their strength (proportion of variance explained),
  universality (how present they are accross landscapes) and consistency
  (how stable they are accross landscapes). At the class level, they
  find seven ``highly universal and consistent structure components''
  (page 700), namely edge contrast, patch shape complexity, aggregation,
  nearest neighbor distance, patch dispersion, large patch dominance and
  neighborhood similarity. At the landscape level, five components are
  identified, namely contagion/diversity (similar to image texture in
  Riitters et al. (Riitters et al. 1995), large patch dominance,
  interspersion/juxtaposition, edge contrast, patch shape variability,
  proximity and nearest neighbor distance. Overall, components are more
  consistent and universal at the landscape level than at the class
  level.
\item
  Uuemaa et al. (Uuemaa et al. 2011) find four main components:
  dominance (also identified by Riitters et al. (Riitters et al. 1995)
  and Cushman et al. (Cushman et al. 2008)), contrast, shape complexity
  and composition.
\item
  Amount (composition) and adjacency (configuration) are the fundamental
  elements of pattern (Riitters 2019)
\item
  Entropy (Nowosad and Stepinski 2019)
\item
  Spatial signatures (Nowosad 2021)
\end{itemize}

\subsubsection{Interpretability}\label{interpretability}

TODO - do landscape metrics measure fragmentation? - normalized metrics
(Wang and Cumming 2011) - (Frazier et al. 2021)

Responses: - incorporating aspects of ecological function into landscape
metrics (e.g., core area, least-cost distances) and moving towards
graph-based and circuit theory approaches (Kupfer 2012). - point-data
analysis (Gustafson 1998) and gradient surface model (Lausch et al.
2015) require fewer assumptions regarding LULC categorization schemes or
spatial scale

\subsection{Surface metrics}\label{surface-metrics}

Surface metrics are based on the gradient surface model using raster
data and continuous values (McGarigal et al. 2009; Cushman et al. 2010)
and were adapted from microscopy and molecular physics (McGarigal et al.
2009; Kedron et al. 2018). The gradient surface model can increase the
resemblance of the data model and the natural world because it allows to
include more heterogeneity within each grid cell (Cushman et al. 2010;
Kedron et al. 2018). Surface metrics are able to quantify various
characteristics, such as roughness, skewness and kurtosis, total and
relative amplitudes, curvatures of local peaks, or surface bearing area
ratios (Cushman et al. 2010). Many surface metrics have analogous
landscape metrics (Cushman et al. 2010), however, using surface metrics
allows to reveal different or additional patterns and potentially
pattern-process links (McGarigal et al. 2009; Borthwick et al. 2020).
Yet, software to calculate these metrics is still rare and further
research is needed into the specific pattern-process links and
ecological meaningful interpretations (Kedron et al. 2018; Frazier
2019).

\subsection{Vector metrics}\label{vector-metrics}

Many landscape ecology studies are based on the raster data model.
However, the decision on the data model is often driven by data and
software availability, or the familiarity with the approach (Costanza et
al. 2019). For example, many LULC maps are provided as raster as they
often related to underlying remote sensing products. Yet, nowadays,
vector data is increasingly available, and most geospatial software
supports it.

Hundreds of metrics have already been developed for raster data, so the
most straightforward approach is to reimplement the same metrics for
vector data. This is possible for most of the metrics, with only several
exceptions of metrics that are specifically related to the raster data
model (Yao et al. 2022). Yet, the vector data model has been widely used
in other fields, such as geography or urban studies. For example, in
urban planning vector metrics are applied to quantify the shapes of
urban areas and characterize the complexity of building footprints
(Basaraner and Cetinkaya 2017). Earlier approaches included four
categories of compactness measures for vector data, namely
perimeter-area, single parameters of related circles, dispersion of
elements of the area around a centroid, and direct comparison to
standard shapes (Maceachren 1985). More recently, a unified theoretical
foundation for measuring shape compactness was introduced using a set of
ten distinct properties of a circle and metrics associated with each of
these properties (Angel et al. 2010)

Vector-based metrics also have their limitations. The most important is
computational complexity, which makes calculations of vector metrics
slower than their raster equivalents. Another technical issue is the
need for the data to be topologically correct, which is often
problematic especially for data from different sources. Last, the
pattern-process link is still less explored for vector-based metrics
compared to raster-based metrics.

\section{Connectivity}\label{connectivity}

Connectivity describes how movement among resources patches is
facilitated or hindered by landscape features such as types, amounts,
and arrangement of LULC (Taylor et al. 1993, 2006). While structural
connectivity focuses on physical and spatial relationships among
resource patches, functional connectivity also includes the movement or
flow of individuals and populations through the landscape (Taylor et al.
2006). Connectivity underpins ecosystem functionality and maintains
biodiversity, and plays a large role in many conservation actions
(Correa Ayram et al. 2016; Hilty et al. 2020). Thus, in the past
multiple approaches how to measure connectivity were developed (Simpkins
et al. 2018; Keeley et al. 2021).

\subsection{Landscape metrics}\label{landscape-metrics-1}

\begin{itemize}
\tightlist
\item
  Many traditional lsm measure connectivity, e.g, ENN
\item
  Recently, also focus in within-patch connectivity
  (https://doi.org/10.1007/s10980-019-00881-0)
\end{itemize}

\subsection{Resistance surfaces}\label{resistance-surfaces}

https://doi.org/10.1111/2041-210X.12750
https://doi.org/10.1007/s10980-019-00789-9
https://doi.org/10.1007/s10980-022-01504-x
https://doi.org/10.1007/s10980-022-01469-x

\subsection{Circuit theory-based}\label{circuit-theory-based}

https://circuitscape.org https://doi.org/10.1111/2041-210X.12689
https://doi.org/10.1016/j.ecolind.2016.03.032

\subsection{Least-cost paths (LCPs)}\label{least-cost-paths-lcps}

https://doi.org/10.1016/S0169-2046(02)00242-6
https://doi.org/10.1007/s40823-016-0006-9

\subsection{Graph theory}\label{graph-theory}

https://doi.org/10.1007/s10980-020-00998-7
https://doi.org/10.1007/s10980-020-01164-9

\subsection{Various}\label{various}

https://doi.org/10.1890/14-1690.1
https://doi.org/10.1016/j.ecolind.2016.03.032
https://doi.org/10.1007/s10980-019-00935-3
https://doi.org/10.1126/science.abg1780

\section{Simulation models}\label{simulation-models}

Simulation models are a powerful tool to study complex adaptive
ecological systems in controllable, reproducible, and replicable
settings (Pascual 2005; Petrovskii and Petrovskaya 2012). Thus, they can
be seen as experimental systems that allow, unlike to the natural world,
all imaginable manipulations in order to advance theoretical
developments or test hypotheses (Peck 2004). Due to the usual temporal
and spatial scales, complex interactions and feedbacks, or scale
mismatches between patterns and processes, simulation models are one of
the major approaches in landscape ecology (Schröder and Seppelt 2006;
Synes et al. 2016). In general, simulations models can be classified
using two major divisions, namely as \emph{i)} predictive or exploratory
models, and \emph{ii)} pattern- or process-based models (Peck 2004;
Synes et al. 2016). Here, we are going to focus mainly on exploratory
models, but include both pattern- and process-based models.

\subsection{Landscapes simulators}\label{landscapes-simulators}

Landscape simulators are typically used as null hypotheses, baselines,
or scenarios because they allow to control certain aspects of the
landscape (Li et al. 2004; Wang and Malanson 2008) and are generally
simpler than ecological simulations models (Hess et al. 2020). Landscape
simulators can be classified into two major categories, namely
pattern-based and process-based approaches (Pe'er et al. 2013;
Langhammer et al. 2019).

Pattern-based approaches simulate landscapes without assuming any
underlying abiotic or biotic processes (i.e., neutral landscape model;
Gardner et al. 1987). The earliest neutral models are based on
percolation theory and assign LULC types randomly to cells in the
landscape(Gardner et al. 1987) or hierarchical models that consider
different spatial scales while assigning cell values (O'Neill et al.
1992). Landscapes characterized by continuous values can be simulated by
fractal models, such as Brownian motion (Palmer 1992). Borrowing from
computer graphics, more recent neutral landscape models make use of
spectral synthesis (e.g., Perlin noise; Etherington 2022) or binary
space partitioning (Etherington et al. 2022). Neutral landscape models
are also able to simulate landscapes dominated by anthropogenic
activities using based on least-cost networks (Etherington et al. 2024).
In order to ensure realistic neutral landscapes several approaches exist
that use comparisons with real landscapes (Inkoom et al. 2017) or are
based on target values optimization (Van Strien et al. 2016;
Justeau-Allaire et al. 2022).

Contrastingly, processes-based approaches explicitly include abiotic or
biotic pattern-forming processes (i.e., landscape generators; Langhammer
et al. 2019; Salecker et al. 2019). Several earlier landscape generators
are based on cellular automata models and are able to simulate urban
growth (Clarke and Gaydos 1998) or deforestation (Soares-Filho et al.
2002). More recent landscape generators allow to simulate patchy
landscapes based on Gibbs processes (Gaucherel et al. 2006),
deforestation based on road and agricultural fields access (Pe'er et al.
2013), agricultural areas dominated by smallholders (Salecker et al.
2019), or vegetation surround watering points in semi-arid savanna
rangelands (combining pattern- and process-bases approaches; Hess et al.
2020).

\subsection{Individual-based models}\label{individual-based-models}

Individual-based models (or agent-based models) simulate discrete
entities that are described by attributes and behaviors and patterns
emerge from bottom-up interactions of individuals with each other and
their environment (DeAngelis and Grimm 2014). In landscape ecology
individual-based models are increasingly used to model especially
social-ecological systems (e.g, Carauta et al. 2018; Dislich et al.
2018; Gonzalez-Redin et al. 2019), but also disturbances (e.g., Seidl
and Rammer 2017; Spies et al. 2017; Kamimura et al. 2019) or
connectivity (e.g., Day et al. 2020). In order to facilitate future
model developments, recent progress includes adaptable modeling
frameworks (Schumaker and Brookes 2018) or reusable building block
(Berger et al. 2024). Furthermore, hybrid approaches that couple or even
integrated different model types with individual-based models could
benefit from rich development history or facilitate corresponding
strengths (O'Sullivan et al. 2016), e.g., linking an individual-based
and ecosystem model (Polhill et al. 2011).

\subsection{Digital twins}\label{digital-twins}

Digital Twins include three connected elements, namely a physical object
or system, its digital representation, and a data exchange between the
former two (Jones et al. 2020; De Koning et al. 2023) and there is an
increasing interest and use of them in both the industry and academia
(Jones et al. 2020). The need for timely and evidence-based decision
making in combination with increasing data availability makes them a
powerful tool for ecology (De Koning et al. 2023). Digital twins are
increasingly applied in agricultural landscapes research, including
topics such as crop and land cultivation, livestock farming, controlled
environment farming, or fertilization management (Cesco et al. 2023;
Purcell et al. 2023). Further applications include planning of rural
ecological landscapes (Tan and Cheng 2024), or exploration of
relationships between urban expansion and vegetation coverage (Zhao et
al. 2022). The use of general data processing, statistical and
mechanistic models, or artificial intelligence allows them to constantly
update the digital representation and analyze, visualize, or predict the
physical counterpart. Thus, Digital Twins can provide real-time state of
nature measurements, early detection of conservation trends,
relationships of ecosystem trends and environmental conditions, impact
assessments of interventions, or identify uncertainties and information
gaps (De Koning et al. 2023).

\section{Software and programming
languages}\label{software-and-programming-languages}

``Open science'' involves full transparency and data access at all
stages of the research processes and integrative and multi-disciplinary
fields such as ecology can massively benefit from this concept (Hampton
et al. 2015). Using open source software is one key aspect of open
science because it ensures transparency and consequently increases trust
in the analyses (Powers and Hampton 2019). Given the increasing data
availability and method complexity, future progress within the research
field will highly depend on available software and analysis tools
(Costanza et al. 2019). In this context, script-based workflows have an
advantage over workflows involving graphical user interfaces because the
former are more shareable and reproducible (Hampton et al. 2015; Powers
and Hampton 2019). Thus, we will focus the following section on
script-based open-source software suitable for landscape ecology.

The open-source, script-based R programming languages (R Core Team 2024)
was first released in 1995 (Smith 2016) with a focus on statistical
computing and data visualization. However, due to its extensions (also
called packages), the language is popular in many academic fields,
including ecology (Lai et al. 2019; Lortie et al. 2020), and it is
suitable for spatial data analysis and landscape ecology (e.g., Bivand
2020; Hesselbarth et al. 2021). Fundamental packages for spatial data
analysis include \emph{sf} for simple-features vector data (Pebesma
2018), \emph{terra} for raster data (Hijmans 2024), \emph{stars} for
spatiotemporal arrays, raster, and vector data cubes (Pebesma and Bivand
2023), and \emph{spatstat} for point pattern data (Baddeley et al.
2015). Several packages allow to create maps, such as \emph{tmap}
(Tennekes 2018) or \emph{ggplot2} (Wickham 2016). For a comprehensive
overview of packages related to geographical spatial data, please see
the CRAN Task View (Bivand and Nowosad 2024). R packages developed
specifically for spatial pattern analysis include, besides others,
\emph{landscapemetrics} (Hesselbarth et al. 2019), \emph{vectormetrics}
{[}still being developed; Matuszek et al. (2024){]}, \emph{motif}
(Nowosad 2021), \emph{rasterdiv} (Rocchini et al. 2021), or
\emph{geodiv} (Smith et al. 2021). Packages with a focus on scale
include e.g., \emph{multilandr} (Huais 2024) or \emph{grainchanger}
(Graham 2024). Connectivity analysis can be conducted using packages
including \emph{lconnect} (Mestre and Silva 2023), or \emph{grainscape}
(Chubaty et al. 2020). Furthermore, neutral landscapes can be simulated
using \emph{NLMR} (Sciaini et al. 2018), and the \emph{graph4lg} (Savary
et al. 2021) provides functions for landscape genetics.

\newpage{}

\subsubsection*{References}\label{references}
\addcontentsline{toc}{subsubsection}{References}

\begingroup
\raggedright

\phantomsection\label{refs}
\begin{CSLReferences}{1}{1}
\bibitem[\citeproctext]{ref-Angel2010}
Angel S, Parent J, Civco DL (2010) Ten compactness properties of
circles: Measuring shape in geography. Canadian Geographies /
G{é}ographies canadiennes 54:441--461.
\url{https://doi.org/10.1111/j.1541-0064.2009.00304.x}

\bibitem[\citeproctext]{ref-Arganaraz2014}
Argañaraz JP, Entraigas I (2014) Scaling functions evaluation for
estimation of landscape metrics at higher resolutions. Ecological
informatics 22:1--12

\bibitem[\citeproctext]{ref-Baddeley2015}
Baddeley A, Rubak E, Turner R (2015) Spatial point patterns: Methodology
and applications with r. Chapman \& Hall/CRC, London, UK

\bibitem[\citeproctext]{ref-Basaraner2017}
Basaraner M, Cetinkaya S (2017) Performance of shape indices and
classification schemes for characterising perceptual shape complexity of
building footprints in GIS. International Journal of Geographical
Information Science 31:1952--1977.
\url{https://doi.org/10.1080/13658816.2017.1346257}

\bibitem[\citeproctext]{ref-Berger2024}
Berger U, Bell A, Barton CM, et al (2024) Towards reusable building
blocks for agent-based modelling and theory development. Environmental
Modelling \& Software 175:106003.
\url{https://doi.org/10.1016/j.envsoft.2024.106003}

\bibitem[\citeproctext]{ref-Bivand2020}
Bivand RS (2020) Progress in the r ecosystem for representing and
handling spatial data. Journal of Geographical Systems.
\url{https://doi.org/10.1007/s10109-020-00336-0}

\bibitem[\citeproctext]{ref-Bivand2024}
Bivand RS, Nowosad J (2024) CRAN task view: Analysis of spatial data.
{https://CRAN.R-project.org/view=Spatial}

\bibitem[\citeproctext]{ref-Borthwick2020}
Borthwick R, de Flamingh A, Hesselbarth MHK, et al (2020) Alternative
quantifications of landscape complementation to model gene flow in
banded longhorn beetles {[}typocerus v. Velutinus (olivier){]}.
Frontiers in Genetics 11:307.
\url{https://doi.org/10.3389/fgene.2020.00307}

\bibitem[\citeproctext]{ref-Carauta2018}
Carauta M, Latynskiy E, Mössinger J, et al (2018) Can preferential
credit programs speed up the adoption of low-carbon agricultural systems
in mato grosso, brazil? Results from bioeconomic microsimulation.
Regional Environmental Change 18:117--128.
\url{https://doi.org/10.1007/s10113-017-1104-x}

\bibitem[\citeproctext]{ref-Cesco2023}
Cesco S, Sambo P, Borin M, et al (2023) Smart agriculture and digital
twins: Applications and challenges in a vision of sustainability.
European Journal of Agronomy 146:126809.
\url{https://doi.org/10.1016/j.eja.2023.126809}

\bibitem[\citeproctext]{ref-Chi2016}
Chi M, Plaza A, Benediktsson JA, et al (2016) Big data for remote
sensing: Challenges and opportunities. Proceedings of the IEEE
104:2207--2219. \url{https://doi.org/10.1109/JPROC.2016.2598228}

\bibitem[\citeproctext]{ref-Chubaty2020a}
Chubaty AM, Galpern P, Doctolero SC (2020) The r toolbox grainscape for
modelling and visualizing landscape connectivity using spatially
explicit networks. Methods in Ecology and Evolution 11:591--595.
\url{https://doi.org/10.1111/2041-210X.13350}

\bibitem[\citeproctext]{ref-Claramunt2012}
Claramunt C (2012)
\href{https://doi.org/10.1007/978-3-642-33999-8_28}{Towards a
spatio-temporal form of entropy}. In: Advances in conceptual modeling.
Springer Berlin Heidelberg, Berlin, Heidelberg, pp 221--230

\bibitem[\citeproctext]{ref-Clarke1998}
Clarke KC, Gaydos LJ (1998) Loose-coupling a cellular automaton model
and GIS: Long-term urban growth prediction for san francisco and
washington/baltimore. International Journal of Geographical Information
Science 12:699--714. \url{https://doi.org/10.1080/136588198241617}

\bibitem[\citeproctext]{ref-CorreaAyram2016}
Correa Ayram CA, Mendoza ME, Etter A, Salicrup DRP (2016) Habitat
connectivity in biodiversity conservation: A review of recent studies
and applications. Progress in Physical Geography: Earth and Environment
40:7--37. \url{https://doi.org/10.1177/0309133315598713}

\bibitem[\citeproctext]{ref-Costanza2019}
Costanza JK, Riitters K, Vogt P, Wickham J (2019) Describing and
analyzing landscape patterns: Where are we now, and where are we going?
Landscape Ecology 34:2049--2055.
\url{https://doi.org/10.1007/s10980-019-00889-6}

\bibitem[\citeproctext]{ref-Cushman2015}
Cushman SA (2015) Thermodynamics in landscape ecology: The importance of
integrating measurement and modeling of landscape entropy. Landscape
Ecology 30:7--10. \url{https://doi.org/10.1007/s10980-014-0108-x}

\bibitem[\citeproctext]{ref-Cushman2016}
Cushman SA (2016) Calculating the configurational entropy of a landscape
mosaic. Landscape Ecology 31:481--489.
\url{https://doi.org/10.1007/s10980-015-0305-2}

\bibitem[\citeproctext]{ref-Cushman2021a}
Cushman SA (2021) Generalizing boltzmann configurational entropy to
surfaces, point patterns and landscape mosaics. Entropy 23:1616.
\url{https://doi.org/10.3390/e23121616}

\bibitem[\citeproctext]{ref-Cushman2010}
Cushman SA, Gutzweiler K, Evans JS, McGarigal K (2010) The gradient
paradigm: A conceptual and analytical framework for landscape ecology.
In: Cushman SA, Huettmann F (eds) Spatial complexity informatics and
wildlife conservation. Springer International Publishing, Basel, CH, pp
83--108

\bibitem[\citeproctext]{ref-Cushman2008Landscape}
Cushman SA, McGarigal K (2008) Landscape metrics, scales of resolution.
In: Designing green landscapes. Springer, pp 33--51

\bibitem[\citeproctext]{ref-Cushman2008Parsimony}
Cushman SA, McGarigal K, Neel MC (2008) Parsimony in landscape metrics:
Strength, universality, and consistency. Ecological indicators
8:691--703

\bibitem[\citeproctext]{ref-Dale1999}
Dale MRT (1999) \href{https://doi.org/10.1017/CBO9780511612589}{Spatial
pattern analysis in plant ecology}, 1st edn. Cambridge University Press

\bibitem[\citeproctext]{ref-Day2020}
Day CC, Zollner PA, Gilbert JH, McCann NP (2020) Individual-based
modeling highlights the importance of mortality and landscape structure
in measures of functional connectivity. Landscape Ecology 35:2191--2208.
\url{https://doi.org/10.1007/s10980-020-01095-5}

\bibitem[\citeproctext]{ref-DeKoning2023}
De Koning K, Broekhuijsen J, Kühn I, et al (2023) Digital twins: Dynamic
model-data fusion for ecology. Trends in Ecology \& Evolution
38:916--926. \url{https://doi.org/10.1016/j.tree.2023.04.010}

\bibitem[\citeproctext]{ref-DeAngelis2014}
DeAngelis DL, Grimm V (2014) Individual-based models in ecology after
four decades. F1000Prime Reports 6:1--6.
\url{https://doi.org/10.12703/P6-39}

\bibitem[\citeproctext]{ref-Dislich2018}
Dislich C, Hettig E, Salecker J, et al (2018) Land-use change in oil
palm dominated tropical landscapes - an agent-based model to explore
ecological and socio-economic trade-offs. PLoS ONE 13:e0190506.
\url{https://doi.org/10.1371/journal.pone.0190506}

\bibitem[\citeproctext]{ref-Dramstad2009}
Dramstad WE (2009) Spatial metrics--useful indicators for society or
mainly fun tools for landscape ecologists? Norsk Geografisk
Tidsskrift-Norwegian Journal of Geography 63:246--254

\bibitem[\citeproctext]{ref-Duncan2023}
Duncan JMA, Boruff B (2023) Monitoring spatial patterns of urban
vegetation: A comparison of contemporary high-resolution datasets.
Landscape and Urban Planning 233:104671.
\url{https://doi.org/10.1016/j.landurbplan.2022.104671}

\bibitem[\citeproctext]{ref-Duque2018}
Duque JC, Laniado H, Polo A (2018) S-maup: Statistical test to measure
the sensitivity to the modifiable areal unit problem. PloS one
13:e0207377

\bibitem[\citeproctext]{ref-Etherington2022}
Etherington TR (2022) Perlin noise as a hierarchical neutral landscape
model. Web Ecology 22:1--6. \url{https://doi.org/10.5194/we-22-1-2022}

\bibitem[\citeproctext]{ref-Etherington2022a}
Etherington TR, Morgan FJ, O'Sullivan D (2022) Binary space partitioning
generates hierarchical and rectilinear neutral landscape models suitable
for human-dominated landscapes. Landscape Ecology 37:1761--1769.
\url{https://doi.org/10.1007/s10980-022-01452-6}

\bibitem[\citeproctext]{ref-Etherington2024}
Etherington TR, O'Sullivan D, Perry GLW, et al (2024) A least-cost
network neutral landscape model of human sites and routes. Landscape
Ecology 39:52. \url{https://doi.org/10.1007/s10980-024-01836-w}

\bibitem[\citeproctext]{ref-Fletcher2023}
Fletcher Jr RJ, Betts MG, Damschen EI, et al (2023) Addressing the
problem of scale that emerges with habitat fragmentation. Global Ecology
and Biogeography 32:828--841

\bibitem[\citeproctext]{ref-Forman1995}
Forman RTT (1995) Land mosaics: The ecology of landscapes and regions.
Cambridge University Press, Cambridge, UK

\bibitem[\citeproctext]{ref-Forman1986}
Forman RTT, Godron M (1986) Landscape ecology. Wiley \& Sons,
Chichester, UK

\bibitem[\citeproctext]{ref-Frazier2019}
Frazier AE (2019) Emerging trajectories for spatial pattern analysis in
landscape ecology. Landscape Ecology 34:2073--2082.
\url{https://doi.org/10.1007/s10980-019-00880-1}

\bibitem[\citeproctext]{ref-Frazier2023}
Frazier AE (2023) Scope and its role in advancing a science of scaling
in landscape ecology. Landscape Ecology 38:637--643

\bibitem[\citeproctext]{ref-Frazier2017}
Frazier AE, Kedron P (2017) Landscape metrics: Past progress and future
directions. Current Landscape Ecology Reports 63--72.
\url{https://doi.org/10.1007/s40823-017-0026-0}

\bibitem[\citeproctext]{ref-Frazier2021}
Frazier AE, Kedron P, Ovando-Montejo GA, Zhao Y (2021) Scaling spatial
pattern metrics: Impacts of composition and configuration on downscaling
accuracy. Landscape Ecology 1--16

\bibitem[\citeproctext]{ref-Gao2019}
Gao P, Li Z (2019) Computation of the boltzmann entropy of a landscape:
A review and a generalization. Landscape Ecology 34:2183--2196.
\url{https://doi.org/10.1007/s10980-019-00814-x}

\bibitem[\citeproctext]{ref-Gao2017}
Gao P, Zhang H, Li Z (2017) A hierarchy-based solution to calculate the
configurational entropy of landscape gradients. Landscape Ecology.
\url{https://doi.org/10.1007/s10980-017-0515-x}

\bibitem[\citeproctext]{ref-Gardner1987}
Gardner RH, Milne BT, Turnei MG, O'Neill RV (1987) Neutral models for
the analysis of broad-scale landscape pattern. Landscape Ecology
1:19--28. \url{https://doi.org/10.1007/BF02275262}

\bibitem[\citeproctext]{ref-Gaucherel2006}
Gaucherel C, Fleury D, Auclair D, Dreyfus P (2006) Neutral models for
patchy landscapes. Ecological Modelling 197:159--170.
\url{https://doi.org/10.1016/j.ecolmodel.2006.02.044}

\bibitem[\citeproctext]{ref-Gergel2017}
Gergel SE, Turner MG (eds) (2017)
\href{https://doi.org/10.1007/978-1-4939-6374-4}{Learning landscape
ecology}. Springer New York, New York, NY

\bibitem[\citeproctext]{ref-Gonzalez-Redin2019}
Gonzalez-Redin J, Gordon IJ, Hill R, et al (2019) Exploring sustainable
land use in forested tropical social-ecological systems: A case-study in
the wet tropics. Journal of Environmental Management 231:940--952.
\url{https://doi.org/10.1016/j.jenvman.2018.10.079}

\bibitem[\citeproctext]{ref-Graham2024}
Graham L (2024) Grainchanger: Moving-window and direct data aggregation.
{https://ropensci.github.io/grainchanger/}

\bibitem[\citeproctext]{ref-Grimm1996}
Grimm V, Frank K, Jeltsch F, et al (1996) Pattern-oriented modelling in
population ecology. The Science of the Total Environment 183:151--166.
\url{https://doi.org/10.1016/0048-9697(95)04966-5}

\bibitem[\citeproctext]{ref-Gustafson1998}
Gustafson EJ (1998) Quantifying landscape spatial pattern: What is the
state of the art? Ecosystems 1:143--156

\bibitem[\citeproctext]{ref-Hampton2015}
Hampton SE, Anderson SS, Bagby SC, et al (2015) The tao of open science
for ecology. Ecosphere 6:1--13.
\url{https://doi.org/10.1890/ES14-00402.1}

\bibitem[\citeproctext]{ref-Hess2020}
Hess B, Dreber N, Liu Y, et al (2020) PioLaG: A piosphere landscape
generator for savanna rangeland modelling. Landscape Ecology
35:2061--2082. \url{https://doi.org/10.1007/s10980-020-01066-w}

\bibitem[\citeproctext]{ref-Hesselbarth2021}
Hesselbarth MHK, Nowosad J, Signer J, Graham LJ (2021) Open-source tools
in r for landscape ecology. Current Landscape Ecology Reports 6:97--111.
\url{https://doi.org/10.1007/s40823-021-00067-y}

\bibitem[\citeproctext]{ref-Hesselbarth2019}
Hesselbarth MHK, Sciaini M, With KA, et al (2019) Landscapemetrics: An
open-source r tool to calculate landscape metrics. Ecography
42:1648--1657. \url{https://doi.org/10.1111/ecog.04617}

\bibitem[\citeproctext]{ref-Hijmans2024}
Hijmans RJ (2024) Terra: Spatial data analysis. R package version
1.7-78. {\(<\)}{https://CRAN.R-project.org/package=terra\(>\)}

\bibitem[\citeproctext]{ref-Hilty2020}
Hilty J, Worboys GL, Keeley A, et al (2020)
\href{https://doi.org/10.2305/IUCN.CH.2020.PAG.30.en}{Guidelines for
conserving connectivity through ecological networks and corridors}.
IUCN, International Union for Conservation of Nature

\bibitem[\citeproctext]{ref-Huais2024}
Huais PY (2024) Multilandr: An r package for multi-scale landscape
analysis. Landscape Ecology 39:140

\bibitem[\citeproctext]{ref-Huckeba2023}
Huckeba G, Andresen B, Roach TNF (2023) Multi-scale spatial ecology
analyses: A kullback information approach. Landscape Ecology
38:645--657. \url{https://doi.org/10.1007/s10980-022-01514-9}

\bibitem[\citeproctext]{ref-Inkoom2017}
Inkoom JN, Frank S, Greve K, Fürst C (2017) Designing neutral landscapes
for data scarce regions in west africa. Ecological Informatics 42:1--13.
\url{https://doi.org/10.1016/j.ecoinf.2017.08.003}

\bibitem[\citeproctext]{ref-Jaric2020}
Jarić I, Correia RA, Brook BW, et al (2020) iEcology: Harnessing large
online resources to generate ecological insights. Trends in Ecology \&
Evolution 35:630--639. \url{https://doi.org/10.1016/j.tree.2020.03.003}

\bibitem[\citeproctext]{ref-Jasiewicz2015}
Jasiewicz J, Netzel P, Stepinski T (2015) GeoPAT: A toolbox for
pattern-based information retrieval from large geospatial databases.
Computers \& Geosciences 80:62--73.
\url{https://doi.org/10.1016/j.cageo.2015.04.002}

\bibitem[\citeproctext]{ref-Jasiewicz2014}
Jasiewicz J, Netzel P, Stepinski TF (2014) Landscape similarity,
retrieval, and machine mapping of physiographic units. Geomorphology
221:104--112. \url{https://doi.org/10.1016/j.geomorph.2014.06.011}

\bibitem[\citeproctext]{ref-Jelinski1996}
Jelinski DE, Wu J (1996) The modifiable areal unit problem and
implications for landscape ecology. Landscape ecology 11:129--140

\bibitem[\citeproctext]{ref-Jones2020}
Jones D, Snider C, Nassehi A, et al (2020) Characterising the digital
twin: A systematic literature review. CIRP Journal of Manufacturing
Science and Technology 29:36--52.
\url{https://doi.org/10.1016/j.cirpj.2020.02.002}

\bibitem[\citeproctext]{ref-Justeau-Allaire2022}
Justeau-Allaire D, Blanchard G, Ibanez T, et al (2022) Fragmented
landscape generator (flsgen): A neutral landscape generator with control
of landscape structure and fragmentation indices. Methods in Ecology and
Evolution 13:1412--1420. \url{https://doi.org/10.1111/2041-210X.13859}

\bibitem[\citeproctext]{ref-Kamimura2019}
Kamimura K, Gardiner B, Dupont S, Finnigan J (2019) Agent-based
modelling of wind damage processes and patterns in forests. Agricultural
and Forest Meteorology 268:279--288.
\url{https://doi.org/10.1016/j.agrformet.2019.01.020}

\bibitem[\citeproctext]{ref-Kedron2018}
Kedron PJ, Frazier AE, Ovando-Montejo GA, Wang J (2018) Surface metrics
for landscape ecology: A comparison of landscape models across
ecoregions and scales. Landscape Ecology 33:1489--1504.
\url{https://doi.org/10.1007/s10980-018-0685-1}

\bibitem[\citeproctext]{ref-Keeley2021}
Keeley ATH, Beier P, Jenness JS (2021) Connectivity metrics for
conservation planning and monitoring. Biological Conservation
255:109008. \url{https://doi.org/10.1016/j.biocon.2021.109008}

\bibitem[\citeproctext]{ref-Kupfer2012}
Kupfer JA (2012) Landscape ecology and biogeography: Rethinking
landscape metrics in a post-FRAGSTATS landscape. Progress in physical
geography 36:400--420

\bibitem[\citeproctext]{ref-Lai2019}
Lai J, Lortie CJ, Muenchen RA, et al (2019) Evaluating the popularity of
r in ecology. Ecosphere 10: \url{https://doi.org/10.1002/ecs2.2567}

\bibitem[\citeproctext]{ref-Langhammer2019}
Langhammer M, Thober J, Lange M, et al (2019) Agricultural landscape
generators for simulation models: A review of existing solutions and an
outline of future directions. Ecological Modelling 393:135--151.
\url{https://doi.org/10.1016/j.ecolmodel.2018.12.010}

\bibitem[\citeproctext]{ref-Lausch2015}
Lausch A, Blaschke T, Haase D, et al (2015) Understanding and
quantifying landscape structure - a review on relevant process
characteristics, data models and landscape metrics. Ecological Modelling
295:31--41. \url{https://doi.org/10.1016/j.ecolmodel.2014.08.018}

\bibitem[\citeproctext]{ref-Lechner2016}
Lechner AM, Rhodes JR (2016) Recent progress on spatial and thematic
resolution in landscape ecology. Current Landscape Ecology Reports
1:98--105. \url{https://doi.org/10.1007/s40823-016-0011-z}

\bibitem[\citeproctext]{ref-Levin1992}
Levin SA (1992) The problem of pattern and scale in ecology: The robert
h. MacArthur award lecture. Ecology 73:1943--1967

\bibitem[\citeproctext]{ref-Li2004b}
Li H, Wu J (2004) Use and misuse of landscape indices. Landscape ecology
19:389--399

\bibitem[\citeproctext]{ref-Li2004a}
Li X, He HS, Wang X, et al (2004) Evaluating the effectiveness of
neutral landscape models to represent a real landscape. Landscape and
Urban Planning 69:137--148.
\url{https://doi.org/10.1016/j.landurbplan.2003.10.037}

\bibitem[\citeproctext]{ref-Lortie2020}
Lortie CJ, Braun J, Filazzola A, Miguel F (2020) A checklist for
choosing between r packages in ecology and evolution. Ecology and
Evolution 10:1098--1105. \url{https://doi.org/10.1002/ece3.5970}

\bibitem[\citeproctext]{ref-Lovelace2019}
Lovelace R, Nowosad J, Münchow J (2019) Geocomputation with r, 1st edn.
Chapman \& Hall/CRC, Boca Raton, USA

\bibitem[\citeproctext]{ref-Maceachren1985}
Maceachren AM (1985) Compactness of geographic shape: Comparison and
evaluation of measures. Geografiska Annaler: Series B, Human Geography
67:53--67. \url{https://doi.org/10.1080/04353684.1985.11879515}

\bibitem[\citeproctext]{ref-Matuszek2024}
Matuszek T, Nowosad J, Sciaini M, et al (2024) Vectormetrics: Landscape
metrics for categorical map patterns in vector data.
{https://github.com/r-spatialecology/vectormetrics}

\bibitem[\citeproctext]{ref-McGarigal2009}
McGarigal K, Tagil S, Cushman SA (2009) Surface metrics: An alternative
to patch metrics for the quantification of landscape structure.
Landscape Ecology 24:433--450.
\url{https://doi.org/10.1007/s10980-009-9327-y}

\bibitem[\citeproctext]{ref-Mestre2023}
Mestre F, Silva B (2023) Lconnect r package: A versatile tool for
evaluating landscape connectivity and prioritizing habitat patches in
conservation research. Ecological Modelling 484:110489.
\url{https://doi.org/10.1016/j.ecolmodel.2023.110489}

\bibitem[\citeproctext]{ref-Nathan2022}
Nathan R, Monk CT, Arlinghaus R, et al (2022) Big-data approaches lead
to an increased understanding of the ecology of animal movement. Science
375:eabg1780. \url{https://doi.org/10.1126/science.abg1780}

\bibitem[\citeproctext]{ref-Nedd2021}
Nedd R, Light K, Owens M, et al (2021) A synthesis of land use/land
cover studies: Definitions, classification systems, meta-studies,
challenges and knowledge gaps on a global landscape. Land 10:994.
\url{https://doi.org/10.3390/land10090994}

\bibitem[\citeproctext]{ref-Newman2019}
Newman EA, Kennedy MC, Falk DA, McKenzie D (2019) Scaling and complexity
in landscape ecology. Frontiers in Ecology and Evolution 7:293.
\url{https://doi.org/10.3389/fevo.2019.00293}

\bibitem[\citeproctext]{ref-Nowosad2021a}
Nowosad J (2021) Motif: An open-source r tool for pattern-based spatial
analysis. Landscape Ecology 36:29--43.
\url{https://doi.org/10.1007/s10980-020-01135-0}

\bibitem[\citeproctext]{ref-Nowosad2019a}
Nowosad J, Stepinski TF (2019) Information theory as a consistent
framework for quantification and classification of landscape patterns.
Landscape Ecology 34:2091--2101.
\url{https://doi.org/10.1007/s10980-019-00830-x}

\bibitem[\citeproctext]{ref-Nowosad2021b}
Nowosad J, Stepinski TF (2021) Pattern-based identification and mapping
of landscape types using multi-thematic data. International Journal of
Geographical Information Science 35:1634--1649.
\url{https://doi.org/10.1080/13658816.2021.1893324}

\bibitem[\citeproctext]{ref-ONeill1992}
O'Neill RV, Gardner RH, Turner MG (1992) A hierarchical neutral model
for landscape analysis. Landscape Ecology 7:55--61.
\url{https://doi.org/10.1007/BF02573957}

\bibitem[\citeproctext]{ref-OSullivan2016}
O'Sullivan D, Evans T, Manson S, et al (2016) Strategic directions for
agent-based modeling: Avoiding the YAAWN syndrome. Journal of Land Use
Science 11:177--187. \url{https://doi.org/10.1080/1747423X.2015.1030463}

\bibitem[\citeproctext]{ref-Openshaw1979}
Openshaw S (1979) A million or so correlated coefficients: Three
experiment on the modifiable areal unit problem. Statistical
applications in the spatial sciences

\bibitem[\citeproctext]{ref-Openshaw1984}
Openshaw S (1984) Ecological fallacies and the analysis of areal census
data. Environment and planning A 16:17--31

\bibitem[\citeproctext]{ref-Palmer1992}
Palmer MW (1992) The coexistence of species in fractal landscapes. The
American Naturalist 139:375--397. \url{https://doi.org/10.1086/285332}

\bibitem[\citeproctext]{ref-Pascual2005}
Pascual M (2005) Computational ecology: From the complex to the simple
and back. PLoS Computional Biology 1:e18.
\url{https://doi.org/10.1371/journal.pcbi.0010018}

\bibitem[\citeproctext]{ref-Peer2013}
Pe'er G, Zurita GA, Schober L, et al (2013) Simple process-based
simulators for generating spatial patterns of habitat loss and
fragmentation: A review and introduction to the g-RaFFe model. PLoS ONE
8:e64968. \url{https://doi.org/10.1371/journal.pone.0064968}

\bibitem[\citeproctext]{ref-Pebesma2018}
Pebesma E (2018) Simple features for r: Standardized support for spatial
vector data. The R Journal 10:439.
\url{https://doi.org/10.32614/RJ-2018-009}

\bibitem[\citeproctext]{ref-Pebesma2023}
Pebesma E, Bivand R (2023)
\href{https://doi.org/10.1201/9780429459016}{Spatial data science: With
applications in r}, 1st edn. {Chapman and Hall/CRC}, New York

\bibitem[\citeproctext]{ref-Peck2004}
Peck SL (2004) Simulation as experiment: A philosophical reassessment
for biological modeling. Trends in Ecology \& Evolution 19:530--534.
\url{https://doi.org/10.1016/j.tree.2004.07.019}

\bibitem[\citeproctext]{ref-Petrovskii2012}
Petrovskii S, Petrovskaya N (2012) Computational ecology as an emerging
science. Interface Focus 2:241--254.
\url{https://doi.org/10.1098/rsfs.2011.0083}

\bibitem[\citeproctext]{ref-Poisot2019}
Poisot T, LaBrie R, Larson E, et al (2019) Data-based, synthesis-driven:
Setting the agenda for computational ecology. Ideas in Ecology and
Evolution 12: \url{https://doi.org/10.24908/iee.2019.12.2.e}

\bibitem[\citeproctext]{ref-Polhill2011}
Polhill JG, Gimona A, Aspinall RJ (2011) Agent-based modelling of land
use effects on ecosystem processes and services. Journal of Land Use
Science 6:75--81. \url{https://doi.org/10.1080/1747423X.2011.558603}

\bibitem[\citeproctext]{ref-Powers2019}
Powers SM, Hampton SE (2019) Open science, reproducibility, and
transparency in ecology. Ecological Applications 29:e01822.
\url{https://doi.org/10.1002/eap.1822}

\bibitem[\citeproctext]{ref-Prlic2012}
Prlić A, Procter JB (2012) Ten simple rules for the open development of
scientific software. PLoS Computational Biology 8:e1002802.
\url{https://doi.org/10.1371/journal.pcbi.1002802}

\bibitem[\citeproctext]{ref-Purcell2023}
Purcell W, Neubauer T, Mallinger K (2023) Digital twins in agriculture:
Challenges and opportunities for environmental sustainability. Current
Opinion in Environmental Sustainability 61:101252.
\url{https://doi.org/10.1016/j.cosust.2022.101252}

\bibitem[\citeproctext]{ref-RCoreTeam2024}
R Core Team (2024) \href{https://www.R-project.org/}{R: A language and
environment for statistical computing}. R Foundation for Statistical
Computing, Vienna, Austria

\bibitem[\citeproctext]{ref-Rao1982}
Rao CR (1982) Diversity and dissimilarity coefficients: A unified
approach. Theoretical Population Biology 21:24--43.
\url{https://doi.org/10.1016/0040-5809(82)90004-1}

\bibitem[\citeproctext]{ref-Remmel2020}
Remmel TK (2020) Distributions of hyper-local configuration elements to
characterize, compare, and assess landscape-level spatial patterns.
Entropy 22:420. \url{https://doi.org/10.3390/e22040420}

\bibitem[\citeproctext]{ref-Renyi1961}
Rényi A (1961) On measures of entropy and information. In: Proceedings
of the fourth berkeley symposium on mathematical statistics and
probability, volume 1: Contributions to the theory of statistics.
University of California Press, pp 547--562

\bibitem[\citeproctext]{ref-Riitters2019}
Riitters K (2019) Pattern metrics for a transdisciplinary landscape
ecology. Landscape Ecology 34:2057--2063

\bibitem[\citeproctext]{ref-Riitters1995}
Riitters KH, O'neill R, Hunsaker C, et al (1995) A factor analysis of
landscape pattern and structure metrics. Landscape ecology 10:23--39

\bibitem[\citeproctext]{ref-Risser1984}
Risser PG, Karr JR, Forman RTT (1984) Landscape ecology: Directions and
approaches. Illinois Natural History Survery Special Publication 2:7--14

\bibitem[\citeproctext]{ref-Rocchini2017}
Rocchini D, Marcantonio M, Ricotta C (2017) Measuring rao's q diversity
index from remote sensing: An open source solution. Ecological
Indicators 72:234--238.
\url{https://doi.org/10.1016/j.ecolind.2016.07.039}

\bibitem[\citeproctext]{ref-Rocchini2021}
Rocchini D, Thouverai E, Marcantonio M, et al (2021) Rasterdiv---an
information theory tailored r package for measuring ecosystem
heterogeneity from space: To the origin and back. Methods in Ecology and
Evolution 12:1093--1102. \url{https://doi.org/10.1111/2041-210X.13583}

\bibitem[\citeproctext]{ref-Salecker2019}
Salecker J, Dislich C, Wiegand K, et al (2019) EFForTS-LGraf: A
landscape generator for creating smallholder-driven land-use mosaics.
PLOS ONE 14:e0222949. \url{https://doi.org/10.1371/journal.pone.0222949}

\bibitem[\citeproctext]{ref-Saura2002}
Saura S (2002) Effects of minimum mapping unit on land cover data
spatial configuration and composition. International Journal of Remote
Sensing 23:4853--4880

\bibitem[\citeproctext]{ref-Saura2007}
Saura S, Castro S (2007) Scaling functions for landscape pattern metrics
derived from remotely sensed data: Are their subpixel estimates really
accurate? ISPRS Journal of Photogrammetry and Remote sensing 62:201--216

\bibitem[\citeproctext]{ref-Saura2001}
Saura S, Martínez-Millán J (2001) Sensitivity of landscape pattern
metrics to map spatial extent. Photogrammetric Engineering \& Remote
Sensing 67:1027--1036

\bibitem[\citeproctext]{ref-Savary2021}
Savary P, Foltête J-C, Moal H, et al (2021) graph4lg: A package for
constructing and analysing graphs for landscape genetics in r. Methods
in Ecology and Evolution 12:539--547.
\url{https://doi.org/10.1111/2041-210X.13530}

\bibitem[\citeproctext]{ref-Schneider2001}
Schneider DC (2001) The rise of the concept of scale in ecology: The
concept of scale is evolving from verbal expression to quantitative
expression. BioScience 51:545--553

\bibitem[\citeproctext]{ref-Schroder2006}
Schröder B, Seppelt R (2006) Analysis of pattern--process interactions
based on landscape models - overview, general concepts, and
methodological issues. Ecological Modelling 199:505--516.
\url{https://doi.org/10.1016/j.ecolmodel.2006.05.036}

\bibitem[\citeproctext]{ref-Schumaker2018}
Schumaker NH, Brookes A (2018) HexSim: A modeling environment for
ecology and conservation. Landscape Ecology 33:197--211.
\url{https://doi.org/10.1007/s10980-017-0605-9}

\bibitem[\citeproctext]{ref-Sciaini2018}
Sciaini M, Fritsch M, Scherer C, Simpkins CE (2018) NLMR and
landscapetools: An integrated environment for simulating and modifying
neutral landscape models in r. Methods in Ecology and Evolution
9:2240--2248. \url{https://doi.org/10.1111/2041-210X.13076}

\bibitem[\citeproctext]{ref-Seidl2017a}
Seidl R, Rammer W (2017) Climate change amplifies the interactions
between wind and bark beetle disturbances in forest landscapes.
Landscape Ecology 32:1485--1498.
\url{https://doi.org/10.1007/s10980-016-0396-4}

\bibitem[\citeproctext]{ref-Shannon1948}
Shannon CE (1948) A mathematical theory of communication. Bell System
Technical Journal 27:379--423.
\url{https://doi.org/10.1002/j.1538-7305.1948.tb01338.x}

\bibitem[\citeproctext]{ref-Simpkins2018}
Simpkins CE, Dennis TE, Etherington TR, Perry GLW (2018) Assessing the
performance of common landscape connectivity metrics using a virtual
ecologist approach. Ecological Modelling 367:13--23.
\url{https://doi.org/10.1016/j.ecolmodel.2017.11.001}

\bibitem[\citeproctext]{ref-Simova2012}
Šı́mová P, Gdulová K (2012) Landscape indices behavior: A review of scale
effects. Applied geography 34:385--394

\bibitem[\citeproctext]{ref-Smith2021b}
Smith AC, Dahlin KM, Record S, et al (2021) The {\textsc{geodiv r}}
package: Tools for calculating gradient surface metrics. Methods in
Ecology and Evolution 12:2094--2100.
\url{https://doi.org/10.1111/2041-210X.13677}

\bibitem[\citeproctext]{ref-Smith2016}
Smith D (2016) Over 16 years of r project history

\bibitem[\citeproctext]{ref-Soares-Filho2002}
Soares-Filho BS, Coutinho Cerqueira G, Lopes Pennachin C (2002)
Dinamica---a stochastic cellular automata model designed to simulate the
landscape dynamics in an amazonian colonization frontier. Ecological
Modelling 154:217--235.
\url{https://doi.org/10.1016/S0304-3800(02)00059-5}

\bibitem[\citeproctext]{ref-Spies2017}
Spies TA, White E, Ager A, et al (2017) Using an agent-based model to
examine forest management outcomes in a fire-prone landscape in oregon,
USA. Ecology and Society 22:art25.
\url{https://doi.org/10.5751/ES-08841-220125}

\bibitem[\citeproctext]{ref-Synes2016}
Synes NW, Brown C, Watts K, et al (2016) Emerging opportunities for
landscape ecological modelling. Current Landscape Ecology Reports
1:146--167

\bibitem[\citeproctext]{ref-Tan2024}
Tan F, Cheng Y (2024) A digital twin framework for innovating rural
ecological landscape control. Environmental Sciences Europe 36:59.
\url{https://doi.org/10.1186/s12302-024-00888-8}

\bibitem[\citeproctext]{ref-Taylor1993}
Taylor PD, Fahrig L, Henein K, Merriam G (1993) Connectivity is a vital
element of landscape structure. Oikos 68:571.
\url{https://doi.org/10.2307/3544927}

\bibitem[\citeproctext]{ref-Taylor2006}
Taylor PD, Fahrig L, With KA (2006)
\href{https://doi.org/10.1017/CBO9780511754821}{Landscape connectivity:
A return to the basics}. In: Crooks KR, Sanjayan M (eds) Connectivity
conservation. Cambridge University Press, Cambridge, pp 29--43

\bibitem[\citeproctext]{ref-Tennekes2018}
Tennekes M (2018) Tmap: Thematic maps in r. Journal of Statistical
Software 84: \url{https://doi.org/10.18637/jss.v084.i06}

\bibitem[\citeproctext]{ref-Turner1989a}
Turner MG (1989) Landscape ecology: The effect of pattern on process.
Annual Review of Ecology and Systematics 20:171--197.
\url{https://doi.org/10.1146/annurev.es.20.110189.001131}

\bibitem[\citeproctext]{ref-Turner1990}
Turner MG (1990) Spatial and temporal analysis of landscape patterns.
Landscape Ecology 4:21--30. \url{https://doi.org/10.1007/BF02573948}

\bibitem[\citeproctext]{ref-Turner2005}
Turner MG (2005) Landscape ecology: What is the state of the science?
Annual Review of Ecology, Evolution, and Systematics 36:319--344.
\url{https://doi.org/10.1146/annurev.ecolsys.36.102003.152614}

\bibitem[\citeproctext]{ref-Turner2015}
Turner MG, Gardner RH (2015) Landscape ecology in theory and practice:
Pattern and process, 2nd edition. Springer, New York

\bibitem[\citeproctext]{ref-Turner1989b}
Turner MG, O'Neill RV, Gardner RH, Milne BT (1989) Effects of changing
spatial scale on the analysis of landscape pattern. Landscape ecology
3:153--162

\bibitem[\citeproctext]{ref-Uuemaa2011}
Uuemaa E, Roosaare J, Oja T, Mander Ü (2011) Analysing the spatial
structure of the estonian landscapes: Which landscape metrics are the
most suitable for comparing different landscapes? Estonian Journal of
Ecology 60:70

\bibitem[\citeproctext]{ref-VanStrien2016}
Van Strien MJ, Slager CTJ, De Vries B, Grêt-Regamey A (2016) An improved
neutral landscape model for recreating real landscapes and generating
landscape series for spatial ecological simulations. Ecology and
Evolution 6:3808--3821. \url{https://doi.org/10.1002/ece3.2145}

\bibitem[\citeproctext]{ref-Vranken2015}
Vranken I, Baudry J, Aubinet M, et al (2015) A review on the use of
entropy in landscape ecology: Heterogeneity, unpredictability, scale
dependence and their links with thermodynamics. Landscape Ecology
30:51--65. \url{https://doi.org/10.1007/s10980-014-0105-0}

\bibitem[\citeproctext]{ref-Wade2003}
Wade TG, Wickham JD, Nash MS, et al (2003) A comparison of vector and
raster GIS methods for calculating landscape metrics used in
environmental assessments. Photogrammetric Engineering \& Remote Sensing
69:1399--1405. \url{https://doi.org/10.14358/PERS.69.12.1399}

\bibitem[\citeproctext]{ref-Wang2008}
Wang Q, Malanson GP (2008) Neutral landscapes: Bases for exploration in
landscape ecology. Geography Compass 2:319--339.
\url{https://doi.org/10.1111/j.1749-8198.2008.00090.x}

\bibitem[\citeproctext]{ref-Wang2011}
Wang X, Cumming SG (2011) Measuring landscape configuration with
normalized metrics. Landscape Ecology 26:723--736

\bibitem[\citeproctext]{ref-Wickham2016}
Wickham H (2016) ggplot2: Elegant graphics for data analysis. Springer
New York, New York, USA

\bibitem[\citeproctext]{ref-Wickham1994}
Wickham JD, Norton DJ (1994) Mapping and analyzing landscape patterns.
Landscape Ecology 9:7--23. \url{https://doi.org/10.1007/BF00135075}

\bibitem[\citeproctext]{ref-Wiegand2003}
Wiegand T, Jeltsch F, Hanski I, Grimm V (2003) Using pattern-oriented
modeling for revealing hidden information: A key for reconciling
ecological theory and application. Oikos 100:209--222.
\url{https://doi.org/10.1034/j.1600-0706.2003.12027.x}

\bibitem[\citeproctext]{ref-Wiens1997}
Wiens JA (1997)
\href{https://doi.org/10.1016/B978-012323445-2/50005-5}{Metapopulation
dynamics and landscape ecology}. In: Metapopulation biology. Elsevier,
pp 43--62

\bibitem[\citeproctext]{ref-Wiersma2022}
Wiersma YF (2022) A review of landscape ecology experiments to
understand ecological processes. Ecological Processes 11:57.
\url{https://doi.org/10.1186/s13717-022-00401-0}

\bibitem[\citeproctext]{ref-Wilson2014}
Wilson G, Aruliah DA, Brown CT, et al (2014) Best practices for
scientific computing. PLoS Biology 12:e1001745.
\url{https://doi.org/10.1371/journal.pbio.1001745}

\bibitem[\citeproctext]{ref-With2019}
With KA (2019) Essentials of landscape ecology, 1st edn. Oxford
University Press, Oxford, UK

\bibitem[\citeproctext]{ref-Wu2013}
Wu J (2013) Key concepts and research topics in landscape ecology
revisited: 30 years after the allerton park workshop. Landscape Ecology
28:1--11. \url{https://doi.org/10.1007/s10980-012-9836-y}

\bibitem[\citeproctext]{ref-Wu2004}
Wu J (2004) Effects of changing scale on landscape pattern analysis:
Scaling relations. Landscape ecology 19:125--138

\bibitem[\citeproctext]{ref-Wu2007}
Wu J (2007) Scale and scaling: A cross-disciplinary perspective. In: Key
topics in landscape ecology. Cambridge University Press, pp 115--142

\bibitem[\citeproctext]{ref-Wu2002b}
Wu J, Hobbs R (2002) Key issues and research priorities in landscape
ecology: An idiosyncratic synthesis. Landscape Ecology 17:355--365.
\url{https://doi.org/10.1023/A:1020561630963}

\bibitem[\citeproctext]{ref-Wu2002a}
Wu J, Shen W, Sun W, Tueller PT (2002) Empirical patterns of the effects
of changing scale on landscape metrics. Landscape ecology 17:761--782

\bibitem[\citeproctext]{ref-Wuest2020}
Wüest RO, Zimmermann NE, Zurell D, et al (2020) Macroecology in the age
of big data -- where to go from here? Journal of Biogeography 47:1--12.
\url{https://doi.org/10.1111/jbi.13633}

\bibitem[\citeproctext]{ref-Yao2022}
Yao Y, Cheng T, Sun Z, et al (2022) VecLI: A framework for calculating
vector landscape indices considering landscape fragmentation.
Environmental Modelling and Software 149:105325.
\url{https://doi.org/10.1016/j.envsoft.2022.105325}

\bibitem[\citeproctext]{ref-Zhang2020}
Zhang H, Wu Z, Lan T, et al (2020) Calculating the wasserstein
metric-based boltzmann entropy of a landscape mosaic. Entropy 22:381.
\url{https://doi.org/10.3390/e22040381}

\bibitem[\citeproctext]{ref-Zhao2022}
Zhao D, Li X, Wang X, et al (2022) Applying digital twins to research
the relationship between urban expansion and vegetation coverage: A case
study of natural preserve. Frontiers in Plant Science 13:840471.
\url{https://doi.org/10.3389/fpls.2022.840471}

\bibitem[\citeproctext]{ref-Zhao2019}
Zhao Y, Zhang X (2019) Calculating spatial configurational entropy of a
landscape mosaic based on the wasserstein metric. Landscape Ecology
34:1849--1858. \url{https://doi.org/10.1007/s10980-019-00876-x}

\end{CSLReferences}

\endgroup

\end{document}