### 2024-08-30 Sections/04-sampling.qmd
Process to connectivity section

---
bibliography: CLER.bib
---

<!-- Person responsible: Bosch -->

# Sampling in landscape ecology

https://doi.org/10.1007/s40823-020-00059-4