### 2024-09-02 Sections/04-sampling.qmd
Update references

---
bibliography: CLER.bib
---

<!-- Person responsible: Bosch -->

# Sampling in landscape ecology

<!-- https://doi.org/10.1007/s40823-020-00059-4 -->