### 2024-09-03 Sections/04-sampling.qmd
Move software to corresponding sections

---
bibliography: CLER.bib
---

<!-- Person responsible: Bosch -->

# Sampling in landscape ecology {#sec-sampling}

<!-- https://doi.org/10.1007/s40823-020-00059-4 -->