### 2024-09-19 Sections/03-connectivity.qmd
Update and shorten connectivity

---
bibliography: CLER.bib
---

<!-- Person responsible: Simpkins, Gerber -->

# Connectivity {#sec-connectivity}

Landscape connectivity describes how landscape features facilitate or impede movement [@Taylor2006LandscapeConnectivity].
It underpins ecosystem functionality, maintains biodiversity and populations, and plays an important role in many conservation actions [@CorreaAyram2016HabitatConnectivity].
Connectivity can be divided into structural and functional connectivity.
Structural connectivity describes the physical arrangement of landscape elements, focusing on spatial aspects such as continuity and adjacency.
It is solely a landscape feature and independent of species' characteristics [@Tischendorf2000UsageMeasurement; @CorreaAyram2016HabitatConnectivity].
Functional connectivity, in contrast, integrates landscape structure with the perceptual, behavioral, and dispersal characteristics of species, making it both species- and landscape-specific [@Adriaensen2003ApplicationLeastcost; @Taylor2006LandscapeConnectivity].

Multiple approaches how to measure connectivity have been developed [@Simpkins2018AssessingPerformance; @Keeley2021ConnectivityMetrics], however, quantifying connectivity has proven both technically and conceptually challenging [@Calabrese2004ComparisonshoppersGuide; @Taylor2006LandscapeConnectivity].
Movement data from many individuals is required to infer the elements in the landscape that individuals preferentially move through [@Simpkins2018AssessingPerformance].
While improving technologies have made tracking animal movement at high tempo-spatial resolution available [@Kays2015TerrestrialAnimal; Nathan2022BigdataApproaches], it is still logistically challenging to track high numbers of individuals.
Thus, in practice connectivity is primarily determined by indirect estimations.
@Calabrese2004ComparisonshoppersGuide distinguished three types of estimates: i) structural connectivity determined by the physical attributes of the landscape, ii) potential connectivity as a combination of physical landscape attributes and limited information about dispersal capabilities of species, and iii) actual connectivity related to observations of individuals moving through a landscape.

### Estimating structural connectivity

Landscape metrics are commonly used to quantify various aspects of landscape structure decomposed into composition and configuration.
These metrics describe structural connectivity in terms of total habitat amount, patch size and isolation, or possible connections between habitat patches.
Additionally, more recently there has been an increasing focus on within-patch connectivity using metrics such as the effective mesh size [@Spanowicz2019MeasuringLandscape].
However, landscape metrics are frequently criticized for an ambiguous link to functional connectivity, however, functionality can be partly included by using ecological meaningful metrics, such as the core area [e.g., @Kupfer2012].

### Estimating functional connectivity

#### Resistance surfaces

Many modern connectivity estimation approaches rely on resistance surfaces to represent the landscape [@Zeller2012EstimatingLandscape].
A resistance surface is a grid-based representation of a landscape where each cell is assigned a value representing the difficulty (or cost) for an individual of a particular species to traverse that cell based on landscape features such as habitat type, topography, or barriers [@Etherington2016LeastcostModelling].
Creating resistance surfaces involves obtaining landscape data for the area of interest along with a sufficiently large buffer area to avoid edge artifacts [@Dutta2022OverviewComputational] and quantifying cost values for each cell using movement data.
This can included expert opinion [widely used due to low effort, however, difficult to measure accuracy\; @Krueger2012RoleExpert], detection data (single point locations of unknown individuals), relocation data (multiple sequential locations of the same individual but at low frequency), pathway data (high-frequency relocation data allowing for movement track inference), or genetic data (samples used to calculate genetic distances between populations).

A large number of software applications have been developed, with these approaches as a base, to estimate connectivity including *ConScape.jl* [@vanMoorter2023AcceleratingAdvances], *Omniscape.jl* [@Landau2021OmniscapejlSoftware], *lconnect* [@Mestre2023LconnectPackage], *grainscape* [@Chubaty2020ToolboxGrainscape], or *Circuitscape* [@Anantharaman2019CircuitscapeJulia].

Least-cost paths assume that individuals have perfect knowledge of the landscape and will choose the path that minimizes resistance costs perfectly with no deviations [@Adriaensen2003ApplicationLeastcost].
By calculating paths between two points, it is possible to estimate connectivity between these points based on the accumulated cost along the path.
This method can also generate accumulated cost surfaces depicting the minimum cost from a single point to all other locations, helping identify reachable areas within a threshold.
The creation of least-cost paths is a well established technique and there are highly optimized and efficient algorithms available for its calculation.

Contrastingly, circuit theory considers all possible pathways between locations simultaneously [@McRae2008UsingCircuit].
In this approach, the landscape is represented as a network of electrical nodes connected by resistors (weighted by the values in the resistance surface), and movement is analogous to electrical current flow.
Circuit theory does not assume that organisms have perfect knowledge of the landscape and is particularly useful when multiple alternative pathways are available.
It can describes isolation by measuring multiple low- or high resistance pathways and identify areas of high movement probability highlighting important corridors and bottlenecks in the landscape [@Cushman2013BiologicalCorridors].

However, resistance surfaces are not free of critique including missing spatial-temporal variability or context dependency [@UnnithanKumar2022ConnectivityModelling; @UnnithanKumar2022MovingLandscape].
Thus, recent developments include the combination with more processes-based modelling approaches [@Schumaker2018HexSimModeling; @Bocedi2021RangeShifter20; @UnnithanKumar2022PathwalkerNew], modelling of dynamic landscape connectivity [@Zeller2020UnderstandingImportance], inclusion of stochasticity and spatial context [@Ribeiro2017LandScapeCorridors], using genetic optimization algorithms [@Peterman2018ResistanceGAPackage], or computational improvements [@Leonard2017GlowSoftware].

#### Graph theory

Following the formative work of @Urban2001LandscapeConnectivity, graph theory has developed into a cornerstone approach to studying landscape connectivity by integrating landscape features and species movement in landscape graphs [@Cushman2013BiologicalCorridors].
Graph theory and connectivity metrics based on it can be applied over various geographical contexts and for various species [@Foltete2020CouplingLandscape], require very little data inputs [@Godet2021ComparisonLandscape], and can provide key information for conservation planning and management.
For example, it can assist wildlife movement corridor planning [e.g., @Wang2023OptimalConfiguration], assist infrastructure placement while minimizing habitat fragmentation [@Pietsch2018ContributionConnectivity], or identify habitat patches that maintain overall landscape connectivity for protection and restoration [@Liu2020UsingLandscape].

Within the landscape delineating the system boundary, discrete habitat patches are modeled as nodes, and the potential movement between patches is modeled as edges.
The edges can be binary, representing whether a connection between two nodes exists or not, or weighted by cost values representing movement efforts from one patch to another [@Petsas2021LandscapeConnectivity].
Connectivity can be quantified through various graph-theoretical metrics describing different aspects, such as the importance of specific patches in maintaining connectivity.
Widely reported connectivity metrics include the probability of connectivity index [@Saura2007NewHabitat], habitat availability metrics quantifying potential and functional connectivity [@Saura2010CommonCurrency; @Pietsch2018ContributionConnectivity], or the integral index of connectivity [@Pascual-Hortal2006ComparisonDevelopment] which is based on a binary connection model [@Saura2010CommonCurrency].

Recently, a new generation of graph-theoretical metrics, commonly reported in combination and addressing caveats of previous metrics [@Hashemi2022ReviewEcological], have been developed to account for landscape features critical for space use and movement of species [@Petsas2021LandscapeConnectivity].
The modeling framework, initially only applied to single species is increasingly being used to model multispecies landscapes [@Clauzel2018CouplingLandscapebased; @Modica2021ImplementationMultispecies; @Liang2023MultispeciesApproach], which account for interspecific movement abilities and habitat preferences.
In addition, modern multiple-layer graph approach offer to model not only spatial but spatio-temporal graphs [@Dale2017SpatiotemporalGraphs; @Uroy2021HowAssess].
Modeling the temporal dynamics helps to identify habitat patches impact connectivity over time, and how these impacts affect biodiversity patterns [@Uroy2021HowAssess].

However, despite the robust theoretical underpinnings of graph theory and its powerful application in landscape connectivity research, contradictory results could emerge due to various ways to construct a graph and different data sources.
In many cases, data availability limits the representation of the modelled landscape, with subsequent implications for the calculated connectivity metrics and inferences of connectivity [Foltete2020CouplingLandscape].
For example, @Godet2021ComparisonLandscape modelled the same landscape using three different data types resulting in different distribution of connectivity values.

Currently, graph-theoretical landscape graphs and connectivity indices can be calculated using several open-source software, including *Conefor* [@Saura2009ConeforSensinode], *Graphab* [@Foltete2012SoftwareTool], the R packages *lconnect* [@Mestre2023LconnectPackage] and *grainscape* [@Chubaty2020ToolboxGrainscape], or Julia’s *ConScape* library [@vanMoorter2023AcceleratingAdvances], among others.