### 2024-09-20 Manuscript.qmd
Update layout

---
title: "Computational methods in landscape ecology"

author:
  - name: Maximilian H.K. Hesselbarth
    affil-id: 1
  - name: Jakub Nowosad
    affil-id: 2,3
  - name: Alida de Flamingh
    affil-id: 4
  - name: Craig E. Simpkins
    affil-id: 5
  - name: Martin Jung
    affil-id: 1
  - name: Gemma Gerber
    affil-id: 1
  - name: Martí Bosch
    affil-id: 6

affiliations:
  - id: 1
    name: International Institute for Applied Systems Analysis, Biodiversity, Ecology, and Conservation Group, Laxenburg, Austria
  - id: 2
    name: Adam Mickiewicz University, Poznań, Poland
  - id: 3
    name: University of Münster, Münster, Germany
  - id: 4
    name: Carl R. Woese Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA 
  - id: 5
    name: Auckland Council, Environmental Services, Auckland, New Zealand
  - id: 6
    name: École polytechnique fédérale de Lausanne, Lausanne, Switzerland

format:
  pdf:
    keep-tex: true
    template-partials:
      - title.tex
    include-in-header:
      text: |
        \usepackage{academicons}
        \usepackage{fontawesome5}
        \usepackage{lineno}\linenumbers
        \usepackage[noblocks]{authblk}
        \usepackage{setspace}
        \renewcommand\Affilfont{\footnotesize}
        \newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid#1}}}
        \newcommand{\mail}[1]{\href{mailto:#1}{\textcolor[HTML]{052E81}{\faEnvelope#1}}}

papersize: a4paper
mainfont: Garamond
fontsize: 11pt
geometry: margin=25mm
indent: 2m
linestretch: 1.5
number-sections: false

bibliography: CLER.bib
csl: springer-basic-author-date.csl
---

\noindent \footnotesize M.H.K. Hesselbarth \orcid{0000-0003-1125-9918} \mail{mhk.hesselbarth@gmail.com}

\noindent \footnotesize J. Nowosad \orcid{0000-0002-1057-3721}

\noindent \footnotesize A. de Flamingh \orcid{0000-0003-1223-6654}

\noindent \footnotesize C.E. Simpkins \orcid{0000-0003-3212-1379}

\noindent \footnotesize M. Jung \orcid{0000-0002-7569-1390}

\noindent \footnotesize G. Gerber \orcid{0000-0002-8722-3113}

\noindent \footnotesize M. Bosch \orcid{0000-0001-8735-9144}

\noindent \small \textit{Keywords: (1) spatial data, (2) spatial patterns, (3) pattern-process link (4) open-source, (5) scientific software}

# Abstract

## Purpose of Review

Landscapes can be defined as mosaics of different land covers, habitats, ecosystems, or land-use systems.
The link between spatial heterogeneity and ecological processes is the core concept in the research field of landscape ecology.
Nowadays, advanced computational methods are essential to the field due to its cross-disciplinary nature, the increasing availability of data, and the complexity of landscape systems.

## Recent findings

This review provides an overview of recent developments in computational methods that have advanced the research field of landscape ecology.
We focus on key topics such as spatial patterns, connectivity, simulations and modelling, landscape genetics, and spatial planning.

## Summary

The review highlights key innovations, challenges, and potential future directions in the field, emphasizing the role of computational methods in addressing complex ecological questions.

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-spatial-patterns.qmd >}}
{{< include Sections/03-connectivity.qmd >}}
{{< include Sections/04-sampling.qmd >}}
{{< include Sections/05-simulations.qmd >}}
{{< include Sections/06-genetics.qmd >}}
{{< include Sections/07-spatial-planning.qmd >}}
{{< include Sections/08-discussion.qmd >}}

# Acknowledgments

We are thankful to the anonymous reviewers and the editors who provided valuable comments to improve the manuscript.

## Funding

No external funding was received.

## Author contributions

All authors contributed particular sections to the manuscript coordinated by MHKH.
Correspondence to [Maximilian H.K. Hesselbarth](mailto:mhk.hesselbarth@gmail.com)

## Conflict of Interest

The authors declare no competing interests.

## Human and Animal Rights and Informed Consent

This article does not contain any studies with human or animal subjects performed by any of the authors.

# References

\footnotesize
\setstretch{1.0}
::: {#refs}
:::